---
title: "PEGASUS new global metabolomics"
subtitle: "Lipidomics positive"
author: "MHuang"
date: "9/22/2020"
output: 
  html_document: 
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

## Packages

pkg <- c("tidyverse", "data.table", "here", "survival", "gtools", "DT")
for (p in pkg) {
        if (require(p, character.only = T)) {
                print(paste0(p, " loaded successfully"))
        } else {
                install.packages(p)
                require(p, character.only = T)
                print(paste0(p, " downloaded and loaded successfully"))
        }
}


## Paths

dat_dir <- "C:/Users/MHuang/Dropbox (Partners HealthCare)/PEGASUS/global_metabolomics_082020/pegasus_global_data_qc/post_qc_data/"
pheno_dir <- "C:/Users/MHuang/Dropbox (Partners HealthCare)/PEGASUS/rpdr_data/rpdr_data_wrangling_analyses/data/"
fig_dir <- "figures/"
res_dir <- "results/"


## Filenames

all_subj_fname <- str_c(pheno_dir, "all_subj_enc_bmi.csv")
med_cs_fname <- str_c(pheno_dir, "med_corticosteroids_summary.csv")
comorb_fname <- str_c(pheno_dir, "comorbidities_of_interest.csv")
lipid_pos_fname <- str_c(dat_dir, "lipid_pos/processed_lipid_pos_halfmin.RData")


## Hard-coded numbers

missing_thld <- 0.30
sig_digits <- 3
alpha_thld <- 0.05
z_95 <- qnorm(p = 0.975)

```

# 1. Load data

```{r}

all_subj <- fread(all_subj_fname)
all_subj[, Collect_Date := as.Date(Collect_Date, format = "%Y-%m-%d")]
all_subj[, .N, .(is.na(EMPI), is.na(Collect_Date))][order(is.na, is.na.1)] %>% 
        setnames(., c("missing_EMPI", "missing_Collect_Date", "N")) %>% print()
all_subj[, asthma_num := case_when(Asthma_Status == TRUE ~ 1, 
                                   Asthma_Status == FALSE ~ 0)]

med_cs <- fread(med_cs_fname)
pheno <- merge(all_subj, med_cs, by = c("EMPI", "SampleGroup"), all.x = T)

comorb <- fread(comorb_fname)
pheno <- merge(pheno, comorb, by = c("EMPI"), all.x = T)


## HILIC-positive halfmin imputed data (outlier checked)

load(lipid_pos_fname, verbose = T)
setnames(samp_dat_final, "Group", "SampleGroup")
lipid_pos_dat <- merge(pheno, samp_dat_final, by = c("sampleID_biobank", "SampleGroup", "Match_ID", 
                                                     "Asthma_Status", "Age", "Gender", "Race", "BMI", "Nonsmoker"))
lipid_pos_mets_info <- copy(mets_info)
lipid_pos_mets_info[grepl("(?i)mouse|cell", tentative_annotation), tentative_annotation := "Unknown"]


## Apply exclusion

unmatched_stratum <- lipid_pos_dat[is.na(EMPI) | is.na(Collect_Date), unique(Stratum)]
lipid_pos_dat[, .N, .(ocs_ard_serum_coll, cortisone_ard_serum_coll, SampleGroup, asthma_num, Asthma_Status)]
asth_to_rm <- lipid_pos_dat[ocs_ard_serum_coll == 1 | is.na(bmi), Biobank_Subject_ID]; length(asth_to_rm)
lipid_pos_dat_trim <- lipid_pos_dat[!(Stratum %in% unmatched_stratum) & 
                                      !(Biobank_Subject_ID %in% asth_to_rm) & !(Match_ID %in% asth_to_rm), ]

lipid_pos_dat_trim[, .N, .(copd_5, a1ad, cf, bronchiectasis, cbronchitis, immunodef, asthma_num)][order(asthma_num)]
comorb_excl <- lipid_pos_dat_trim[copd_5 == 1 | a1ad == 1 | cf == 1 | bronchiectasis == 1 | cbronchitis == 1 | 
                                    immunodef == 1, Biobank_Subject_ID]; length(comorb_excl)
lipid_pos_dat_trim2 <- lipid_pos_dat_trim[!(Biobank_Subject_ID %in% comorb_excl) & !(Match_ID %in% comorb_excl), ]


## Subgroup by ICS use in 1 year before collection

lipid_pos_noics_1y <- lipid_pos_dat_trim[Stratum %in% lipid_pos_dat[SampleGroup == "Asthmatic" & ics_trim_totnum_1y == 0, Stratum], ]
lipid_pos_ics_1y <- lipid_pos_dat_trim[Stratum %in% lipid_pos_dat[SampleGroup == "Asthmatic" & ics_trim_totnum_1y >= 1, Stratum], ]


## Subgroup by OCS use in 1 year before collection

lipid_pos_noocs_1y <- lipid_pos_dat_trim[Stratum %in% lipid_pos_dat[SampleGroup == "Asthmatic" & ocs_trim_totnum_1y == 0, Stratum], ]
lipid_pos_ocs_1y <- lipid_pos_dat_trim[Stratum %in% lipid_pos_dat[SampleGroup == "Asthmatic" & ocs_trim_totnum_1y >= 1, Stratum], ]


## Remove original metabolomic data files with non-specific names

rm(mets_info, samp_dat_final, samp_dat_origscale, samp_na, samp_mets_cols)

```

# 2. Conditional logistic models

On features that passed QC with <= 30% missing

```{r}

feat_list <- lipid_pos_mets_info[is.na(qc_rm) & pct_na_met <= missing_thld, feature]; length(feat_list)

clogi_res_fnc <- function(outc, mets_list, dat, mets_info, add_covar = "", 
                          first_cols = c("feature", "beta", "pval", "fdr_bh", "or", "lower95", "upper95", 
                                         "tentative_annotation", "totaln", "casen", "rt_min", "mz", "cv", "pct_na_met"), 
                          na_thld = missing_thld) {
        
        clogi_mdl_tmp <- function(met) {
                
                fit <- eval(parse(text = str_c("clogit(", outc, " ~ ", met, " + strata(Stratum)", 
                                               add_covar, ", data = ", dat, ")")))
                
                coef <- summary(fit)$coefficients
                coef <- coef[rownames(coef) == met, ]
                
                confint <- summary(fit)$conf.int
                confint <- confint[rownames(confint) == met, -c(1:2)]
                
                totaln <- summary(fit)$n
                casen <- summary(fit)$nevent
                
                output <- t(c(outc, met, coef, confint, totaln, casen)) %>% as.data.table()
                colnames(output) <- c("outc", "feature", "beta", "or", "se", "zval", "pval", "lower95", "upper95", "totaln", "casen")
                output[, (colnames(output)[-c(1:2)]) := lapply(.SD, as.numeric), .SDcols = colnames(output)[-c(1:2)]]
                
                return(output)
        }
        
        res <- lapply(mets_list, function(x) {clogi_mdl_tmp(met = x)}) %>% rbindlist()
        
        eval(parse(text = str_c("mets_info <- copy(", mets_info, ")")))
        res <- merge(res, 
                     mets_info[, .(feature, rt_min, mz, tentative_annotation, cv, pct_na_met, skewness_post)], 
                     by = "feature")
        res[pct_na_met <= na_thld, fdr_bh := p.adjust(pval, method = "BH")] 
        
        setcolorder(res, first_cols)
        res <- res[order(pval)]
        return(res)
}

```

# 3. Analyses accounting for matched strata
## 3.1. Adjusted for BMI

Not excluding COPD, cystic fibrosis, bronchiectasis, chronic bronchitis, alpha-1-antitrypsin deficiency, and immune deficiencies

```{r}

res_3_1 <- clogi_res_fnc(outc = "asthma_num", 
                         mets_list = feat_list, 
                         dat = "lipid_pos_dat_trim", 
                         mets_info = "lipid_pos_mets_info", 
                         add_covar = " + bmi")
res_3_1[pval < alpha_thld, ] %>% nrow() 
res_3_1[fdr_bh < alpha_thld, ] %>% nrow() 
res_3_1[pval < alpha_thld/nrow(res_3_1), ] %>% nrow() 
datatable(res_3_1[tentative_annotation != "Unknown"], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)
datatable(res_3_1, filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)

```

## 3.2. Adjusted for BMI & excluding certain comorbities

Excluding COPD, cystic fibrosis, bronchiectasis, chronic bronchitis, alpha-1-antitrypsin deficiency, and immune deficiencies

```{r}

res_3_2 <- clogi_res_fnc(outc = "asthma_num", 
                         mets_list = feat_list, 
                         dat = "lipid_pos_dat_trim2", 
                         mets_info = "lipid_pos_mets_info", 
                         add_covar = " + bmi")
res_3_2[pval < alpha_thld, ] %>% nrow() 
res_3_2[fdr_bh < alpha_thld, ] %>% nrow() 
res_3_2[pval < alpha_thld/nrow(res_3_2), ] %>% nrow() 
datatable(res_3_2[tentative_annotation != "Unknown"], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)
datatable(res_3_2, filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)

```

## 3.3. Adjusted for BMI and certain comorbidities & excluding certain comorbities

Excluding COPD, cystic fibrosis, bronchiectasis, chronic bronchitis, alpha-1-antitrypsin deficiency, and immune deficiencies
Adjusting for pulmonary hypertension, allergic rhinitis, congestive heart failure, gerd

```{r}

res_3_3 <- clogi_res_fnc(outc = "asthma_num", 
                         mets_list = feat_list, 
                         dat = "lipid_pos_dat_trim2", 
                         mets_info = "lipid_pos_mets_info", 
                         add_covar = " + bmi + pulhtn + arhinitis + chf + gerd")
res_3_3[pval < alpha_thld, ] %>% nrow() 
res_3_3[fdr_bh < alpha_thld, ] %>% nrow() 
res_3_3[pval < alpha_thld/nrow(res_3_3), ] %>% nrow() 
datatable(res_3_3[tentative_annotation != "Unknown"], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)
datatable(res_3_3, filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "pct_na_met", "cv"), digits = 3)

```

# 4. Save results

```{r}

res_lipid_pos <- copy(res_3_2)
save(lipid_pos_dat, lipid_pos_dat_trim, lipid_pos_dat_trim2, lipid_pos_mets_info, res_lipid_pos, 
     file = here(res_dir, "res_lipid_pos_clogi.RData"))

```

# 5. Session info

```{r}

sessionInfo()

```
