---
title: "PEGASUS Analyses Accounting for Medications"
subtitle: "Targeted sphingolipids"
author: "MHuang"
date: "8/18/2020"
output: 
  html_document: 
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

## Packages

pkg <- c("tidyverse", "data.table", "here", "DT", "tableone", "survival", "gtools", 
         "ggrepel", "VennDiagram", "RColorBrewer")
for (p in pkg) {
        if (require(p, character.only = T)) {
                print(paste0(p, " loaded successfully"))
        } else {
                install.packages(p)
                require(p, character.only = T)
                print(paste0(p, " downloaded and loaded successfully"))
        }
}


## Paths

dat_dir <- "data/"
res_dir <- "results/"
fig_dir <- "figures/"
sph_dat_dir <- "C:/Users/MHuang/Dropbox (Partners HealthCare)/PEGASUS/sphingolipids_targeted/data/"


## Filenames

all_subj_fname <- here(dat_dir, "all_subj_enc_bmi.csv")
med_cs_smry_fname <- here(dat_dir, "med_corticosteroids_summary.csv")
asth_exac_smry_fname <- here(dat_dir, "asthma_exacerbation_summary.csv")
wbc_lastbf_fname <- here(dat_dir, "wbc_count_last_before_serum_collection.csv")
wbc_closest_fname <- here(dat_dir, "wbc_count_closest_to_serum_collection.csv")
comorb_fname <- here(dat_dir, "comorbidities_of_interest.csv")
sph_dat_fname <- str_c(sph_dat_dir, "eda_updated.RData")


## Hard-coded numbers

missing_thld <- 0.30
sig_digits <- 3
alpha_thld <- 0.05
z_95 <- qnorm(p = 0.975)

```

# 1. Sphingolipids-WAS - conditional logistic
## 1.1. Read in data & apply exclusion

```{r}

## Phenotypic data 

all_subj <- fread(all_subj_fname)
all_subj[, Collect_Date := as.Date(Collect_Date, format = "%Y-%m-%d")]
summary(all_subj[, Collect_Date])
all_subj[, .N, .(is.na(EMPI), is.na(Collect_Date))][order(is.na, is.na.1)] %>% 
        setnames(., c("missing_EMPI", "missing_Collect_Date", "N")) %>% print()
ggplot(all_subj[!is.na(Collect_Date), ]) + 
        geom_histogram(aes(Collect_Date), bins = 100)

med_cs_smry <- fread(med_cs_smry_fname)
asth_exac_smry <- fread(asth_exac_smry_fname)
# wbc_lastbf <- fread(wbc_lastbf_fname)
wbc_closest <- fread(wbc_closest_fname)

pheno <- merge(all_subj[!(Stratum %in% all_subj[is.na(EMPI) | is.na(Collect_Date), Stratum]), ], 
               med_cs_smry, by = c("EMPI", "SampleGroup"), all.x = T)
pheno <- merge(pheno, asth_exac_smry, by = c("EMPI", "SampleGroup"), all.x = T)
# pheno <- merge(pheno, wbc_lastbf, by = c("EMPI"), all.x = T)
pheno <- merge(pheno, wbc_closest, by = c("EMPI"), all.x = T)


## Sphingolipids data

load(sph_dat_fname, verbose = T)
sph <- copy(mets_pheno)
setnames(sph, "SubjectID", "Biobank_Subject_ID")
sph_info <- copy(mets_info)
rm(mets_pheno, mets_info)


## Add a variable for subgroup of sphingolipid

sph_info[, sph_subgrp := case_when(grepl("^Cer|DhCer", metabolite) ~ "Ceramide/Cer1P/DhCer", 
                                   grepl("^HexCer|LacCer", metabolite) ~ "Hex/LacCeramide", 
                                   metabolite %in% sm_list ~ "Sphingomyelin", 
                                   grepl("^S1P|Sph", metabolite) ~ "S1P/sphingosine", 
                                   grepl("^Spa1P|Spad", metabolite) ~ "Spa1P/Sphinganine", 
                                   TRUE ~ "Other sphingolipid")]


## Merge sphingolipid data with phenotypic data

sph_pheno <- merge(pheno, 
                   subset(sph, select = c("Biobank_Subject_ID", "Comments_sample", sph_list, 
                                          "redness", "milkiness")), 
                   by = "Biobank_Subject_ID")

sph_pheno[, (sph_list) := lapply(.SD, function(x) scale(x, center = T, scale = T)), .SDcols = sph_list]

sph_pheno[, .N, Comments_sample]

```

## 1.2. Model and plot functions

```{r}

clogi_res_fnc <- function(outc, mets_list, dat, add_covar = "", 
                          first_cols = c("metabolite", "beta", "pval", "fdr_bh", "or", "lower95", "upper95", 
                                         "totaln", "casen", "pct_na_met")) {
        
        clogi_mdl_tmp <- function(met) {
                
                fit <- eval(parse(text = str_c("clogit(", outc, " ~ ", met, " + strata(Stratum)", 
                                               add_covar, ", data = ", dat, ")")))
                
                coef <- summary(fit)$coefficients
                coef <- coef[rownames(coef) == met, ]
                
                confint <- summary(fit)$conf.int
                confint <- confint[rownames(confint) == met, -c(1:2)]
                
                totaln <- summary(fit)$n
                casen <- summary(fit)$nevent
                
                output <- t(c(outc, met, coef, confint, totaln, casen)) %>% as.data.table()
                colnames(output) <- c("outc", "metabolite", "beta", "or", "se", "zval", "pval", 
                                      "lower95", "upper95", "totaln", "casen")
                output[, (colnames(output)[-c(1:2)]) := lapply(.SD, as.numeric), .SDcols = colnames(output)[-c(1:2)]]
                
                return(output)
        }
        
        res <- lapply(mets_list, function(x) {clogi_mdl_tmp(met = x)}) %>% rbindlist()
        res <- merge(res, sph_info, by = "metabolite")
        res[, fdr_bh := p.adjust(pval, method = "BH")] 
        
        setcolorder(res, first_cols)
        res <- res[order(pval)]
        return(res)
}

volcano_plot_fnc <- function(res, title = NULL, subtitle = NULL, sig_cutoff = c("pval", "fdr_bh")) {
        ggplot(res, aes(x = beta, y = -log10(pval))) + 
                geom_point(aes(color = sph_subgrp), alpha = 0.6, size = 3) +
                scale_color_manual(values = c("#007DBA", "#A86F5F", "#E34F33", "#00AE91", "#F8A3BC", "#FFC658"), 
                                   name = "Sphingolipid subgroup") + 
                labs(title = title, 
                     subtitle = subtitle, 
                     x = expression('Estimated '*beta), 
                     y = expression('-log'[10]*'('*italic(P)*'-value)')) + 
                geom_label_repel(aes(label = eval(parse(text = str_c("ifelse(", sig_cutoff, 
                                                                     " < alpha_thld, metabolite, '')")))),
                                 box.padding   = 0.25,
                                 point.padding = 0.5,
                                 segment.color = 'grey50') + 
                theme_minimal() + 
                theme(title = element_text(size = 12), 
                      axis.title = element_text(size = 12), 
                      axis.text = element_text(size = 12), 
                      legend.text = element_text(size = 12), 
                      legend.justification = "bottom")
}

```

## 1.3. Overall asthmatics vs controls

```{r}

# volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, 
#                                      dat = "sph_pheno"), 
#                  title = "Asthmatics vs controls (conditional logistic models)", 
#                  sig_cutoff = "pval")

volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, 
                                     dat = "sph_pheno", add_covar = " + bmi"), 
                 title = "Asthmatics vs controls (conditional logistic models)", 
                 subtitle = "adjusted for BMI closest to serum collection", 
                 sig_cutoff = "pval")
fwrite(clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, 
                     dat = "sph_pheno", add_covar = " + bmi"), 
       file = here(res_dir, "sphingolipids_clogi_results_bmi_adjusted.csv"))

volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, 
                                     dat = "sph_pheno[Stratum %in% sph_pheno[ocs_ard_serum_coll == 0, Stratum], ]", 
                                     add_covar = " + bmi"), 
                 title = "Asthmatics vs controls (conditional logistic models)", 
                 subtitle = "adjusted for BMI closest to serum collection; removed OCS w/in 10d", 
                 sig_cutoff = "pval")
fwrite(clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, 
                     dat = "sph_pheno[Stratum %in% sph_pheno[ocs_ard_serum_coll == 0, Stratum], ]", 
                     add_covar = " + bmi"), 
       file = here(res_dir, "sphingolipids_clogi_results_bmi_adjusted_ocs_in_10d_removed.csv"))

```

## 1.4. Subgroup by ICS prescription in asthmatics

```{r}

sph_pheno[SampleGroup == "Asthmatic", .N, .(ics_trim_totnum_1y == 0)]
noics_1y <- sph_pheno[Stratum %in% sph_pheno[SampleGroup == "Asthmatic" & ics_trim_totnum_1y == 0, Stratum], ]
ics_1y <- sph_pheno[Stratum %in% sph_pheno[SampleGroup == "Asthmatic" & ics_trim_totnum_1y >= 1, Stratum], ]

```

### 1.4.1. Asthmatics with no ICS prescription w/in 1yr before serum collection vs controls

```{r}

datatable(noics_1y[is.na(bmi), .(Stratum, Asthma_Status)][order(Stratum, Asthma_Status)])

# volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "noics_1y"), 
#                  title = "Asthmatics with no ICS prescription w/in 1yr before serum collection vs controls", 
#                  sig_cutoff = "pval")

volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "noics_1y", 
                                     add_covar = " + bmi"), 
                 title = "Asthmatics with no ICS prescription w/in 1yr before serum collection vs controls", 
                 subtitle = "conditional logistic adjusted for BMI closest to serum collection", 
                 sig_cutoff = "pval")

volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, 
                                     dat = "noics_1y[Stratum %in% sph_pheno[ocs_ard_serum_coll == 0, Stratum], ]", 
                                     add_covar = " + bmi"), 
                 title = "Asthmatics with no ICS prescription w/in 1yr before serum collection vs controls", 
                 subtitle = "conditional logistic adjusted for BMI closest to serum collection; removed OCS w/in 10d", 
                 sig_cutoff = "pval")

```

### 1.4.2. Asthmatics with >=1 ICS prescription w/in 1yr before serum collection vs controls

```{r}

datatable(ics_1y[is.na(bmi), .(Stratum, Asthma_Status)][order(Stratum, Asthma_Status)])

# volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "ics_1y"), 
#                  title = "Asthmatics with >=1 ICS prescription w/in 1yr before serum collection vs controls", 
#                  sig_cutoff = "pval")

volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "ics_1y", 
                                     add_covar = " + bmi"), 
                 title = "Asthmatics with >=1 ICS prescription w/in 1yr before serum collection vs controls", 
                 subtitle = "conditional logistic adjusted for BMI closest to serum collection", 
                 sig_cutoff = "pval")

volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, 
                                     dat = "ics_1y[Stratum %in% sph_pheno[ocs_ard_serum_coll == 0, Stratum], ]", 
                                     add_covar = " + bmi"), 
                 title = "Asthmatics with >=1 ICS prescription w/in 1yr before serum collection vs controls", 
                 subtitle = "conditional logistic adjusted for BMI closest to serum collection; removed OCS w/in 10d", 
                 sig_cutoff = "pval")

```

## 1.5. Subgroup by OCS prescription in asthmatics

```{r}

sph_pheno[SampleGroup == "Asthmatic", .N, .(ocs_trim_totnum_1y == 0)]
noocs_1y <- sph_pheno[Stratum %in% sph_pheno[SampleGroup == "Asthmatic" & ocs_trim_totnum_1y == 0, Stratum], ]
ocs_1y <- sph_pheno[Stratum %in% sph_pheno[SampleGroup == "Asthmatic" & ocs_trim_totnum_1y >= 1, Stratum], ]

```

### 1.5.1. Asthmatics with no OCS prescription w/in 1yr before serum collection vs controls

```{r}

datatable(noocs_1y[is.na(bmi), .(Stratum, Asthma_Status)][order(Stratum, Asthma_Status)])

volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "noocs_1y", 
                                     add_covar = " + bmi"), 
                 title = "Asthmatocs with no OCS prescription w/in 1yr before serum collection vs controls", 
                 subtitle = "conditional logistic adjusted for BMI closest to serum collection", 
                 sig_cutoff = "pval")

```

### 1.5.2. Asthmatics with >=1 OCS prescription w/in 1yr before serum collection vs controls

```{r}

datatable(ocs_1y[is.na(bmi), .(Stratum, Asthma_Status)][order(Stratum, Asthma_Status)])

volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "ocs_1y", 
                                     add_covar = " + bmi"), 
                 title = "Asthmatocs with >=1 OCS prescription w/in 1yr before serum collection vs controls", 
                 subtitle = "conditional logistic adjusted for BMI closest to serum collection", 
                 sig_cutoff = "pval")

```

## 1.6. Subgroup by exacerbation count in asthmatics

```{r}

sph_pheno[SampleGroup == "Asthmatic", .N, .(asthma_exacerbation_trim_totnum_1y == 0)]
noexac_1y <- sph_pheno[Stratum %in% sph_pheno[SampleGroup == "Asthmatic" & 
                                                asthma_exacerbation_trim_totnum_1y == 0, Stratum], ]
exac_1y <- sph_pheno[Stratum %in% sph_pheno[SampleGroup == "Asthmatic" & 
                                              asthma_exacerbation_trim_totnum_1y >= 1, Stratum], ]

```

### 1.6.1. Asthmatics with no exacerbation w/in 1yr before serum collection vs controls

```{r}

datatable(noexac_1y[is.na(bmi), .(Stratum, Asthma_Status)][order(Stratum, Asthma_Status)])

volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "noexac_1y", 
                                     add_covar = " + bmi"), 
                 title = "Asthmatexac with no exacerbation w/in 1yr before serum collection vs controls", 
                 subtitle = "conditional logistic adjusted for BMI closest to serum collection", 
                 sig_cutoff = "pval")

```

### 1.6.2. Asthmatics with >=1 exacerbation w/in 1yr before serum collection vs controls

```{r}

datatable(exac_1y[is.na(bmi), .(Stratum, Asthma_Status)][order(Stratum, Asthma_Status)])

volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "exac_1y", 
                                     add_covar = " + bmi"), 
                 title = "Asthmatexac with >=1 exacerbation w/in 1yr before serum collection vs controls", 
                 subtitle = "conditional logistic adjusted for BMI closest to serum collection", 
                 sig_cutoff = "pval")

```

## 1.7. Subgroup by eosinophil count high/low in asthmatics

```{r}

sph_pheno[SampleGroup == "Asthmatic", .N, .(eos >= 0.3)]
eos_high <- sph_pheno[Stratum %in% sph_pheno[SampleGroup == "Asthmatic" & eos >= 0.3, Stratum], ]
eos_low <- sph_pheno[Stratum %in% sph_pheno[SampleGroup == "Asthmatic" & eos < 0.3, Stratum], ]

```

### 1.7.1. Asthmatics with eosinophil >= 0.3 K/uL vs controls

```{r}

datatable(eos_high[is.na(bmi), .(Stratum, Asthma_Status)][order(Stratum, Asthma_Status)])

volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "eos_high", 
                                     add_covar = " + bmi"), 
                 title = "Asthmatics with eosinophil >= 0.3 K/uL vs controls", 
                 subtitle = "conditional logistic adjusted for BMI closest to serum collection", 
                 sig_cutoff = "pval")

```

### 1.7.2. Asthmatics with eosinophil < 0.3 K/uL vs controls

```{r}

datatable(eos_low[is.na(bmi), .(Stratum, Asthma_Status)][order(Stratum, Asthma_Status)])

volcano_plot_fnc(res = clogi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "eos_low", 
                                     add_covar = " + bmi"), 
                 title = "Asthmatics with eosinophil < 0.3 K/uL vs controls", 
                 subtitle = "conditional logistic adjusted for BMI closest to serum collection", 
                 sig_cutoff = "pval")

```

# 2. Sphingolipids-WAS - ordinary logistic
## 2.1. Model function

```{r}

logi_res_fnc <- function(outc, mets_list, dat, add_covar = "", 
                         first_cols = c("metabolite", "beta", "pval", "fdr_bh", "or", 
                                        "lower95", "upper95", "pct_na_met")) {
        
        logi_mdl_tmp <- function(met) {
                
                fit <- eval(parse(text = str_c("glm(", outc, " ~ ", met, " + Age + Gender + Race", 
                                               add_covar, ", data = ", dat, ", family = 'binomial')")))
                
                coef <- summary(fit)$coefficients
                coef <- coef[rownames(coef) == met, ]
                
                or <- exp(coef[1])
                
                suppressMessages(confint <- confint(fit))
                confint <- confint[rownames(confint) == met, ]
                
                output <- t(c(outc, met, coef, or, confint)) %>% as.data.table()
                colnames(output) <- c("outc", "metabolite", "beta", "se", "zval", "pval", "or", 
                                      "lower95", "upper95")
                output[, (colnames(output)[-c(1:2)]) := lapply(.SD, as.numeric), .SDcols = colnames(output)[-c(1:2)]]
                
                return(output)
        }
        
        res <- lapply(mets_list, function(x) {logi_mdl_tmp(met = x)}) %>% rbindlist()
        res <- merge(res, sph_info, by = "metabolite")
        res[, fdr_bh := p.adjust(pval, method = "BH")] 
        
        setcolorder(res, first_cols)
        res <- res[order(pval)]
        return(res)
}

```

## 2.2. Overall asthmatics vs controls

```{r}

volcano_plot_fnc(res = logi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, 
                                    dat = "sph_pheno", add_covar = " + bmi"), 
                 title = "Asthmatics vs controls (ordinary logistic models)", 
                 subtitle = "adjusted for age, gender, race, BMI closest to serum collection", 
                 sig_cutoff = "pval")

```

## 2.3. Subgroup by ICS prescription in asthmatics
### 2.3.1. Asthmatics with no ICS prescription w/in 1yr before serum collection vs controls

```{r}

volcano_plot_fnc(res = logi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "noics_1y", 
                                    add_covar = " + bmi"), 
                 title = "Asthmatics with no ICS prescription w/in 1yr before serum collection vs controls", 
                 subtitle = "ordinary logistic adjusted for age, gender, race, BMI closest to serum collection", 
                 sig_cutoff = "pval")

```

### 2.3.2. Asthmatics with >=1 ICS prescription w/in 1yr before serum collection vs controls

```{r}

volcano_plot_fnc(res = logi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "ics_1y", 
                                    add_covar = " + bmi"), 
                 title = "Asthmatics with >=1 ICS prescription w/in 1yr before serum collection vs controls", 
                 subtitle = "ordinary logistic adjusted for age, gender, race, BMI closest to serum collection", 
                 sig_cutoff = "pval")

```

## 2.4. Subgroup by OCS prescription in asthmatics
### 2.4.1. Asthmatics with no exacerbation w/in 1yr before serum collection vs controls

```{r}

volcano_plot_fnc(res = logi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "noexac_1y", 
                                    add_covar = " + bmi"), 
                 title = "Asthmatexac with no exacerbation w/in 1yr before serum collection vs controls", 
                 subtitle = "ordinary logistic adjusted for age, gender, race, BMI closest to serum collection", 
                 sig_cutoff = "pval")

```

### 2.4.2. Asthmatics with >=1 exacerbation w/in 1yr before serum collection vs controls

```{r}

volcano_plot_fnc(res = logi_res_fnc(outc = "Asthma_Status", mets_list = sph_list, dat = "exac_1y", 
                                    add_covar = " + bmi"), 
                 title = "Asthmatexac with >=1 exacerbation w/in 1yr before serum collection vs controls", 
                 subtitle = "ordinary logistic adjusted for age, gender, race, BMI closest to serum collection", 
                 sig_cutoff = "pval")

```

# 3. Within asthmatics - corticosteroids medication & metabolites
## 3.1. Model function

```{r}

linear_res_fnc <- function(expo, mets_list, dat, add_covar = "", 
                           first_cols = c("metabolite", "beta", "pval", "fdr_bh", "pct_na_met")) {
        
        linear_mdl_tmp <- function(met) {
                
                fit <- eval(parse(text = str_c("lm(", met, " ~ ", expo, " + Age + Gender + Race", 
                                               add_covar, ", data = ", dat, ")")))
                
                coef <- summary(fit)$coefficients
                coef <- coef[rownames(coef) == expo, ]

                # totaln <- summary(fit)$fstatistic[2] + summary(fit)$fstatistic[3] + 2
                
                output <- t(c(expo, met, coef)) %>% as.data.table()
                colnames(output) <- c("expo", "metabolite", "beta", "se", "tval", "pval")
                output[, (colnames(output)[-c(1:2)]) := lapply(.SD, as.numeric), .SDcols = colnames(output)[-c(1:2)]]
                
                return(output)
        }
        
        res <- lapply(mets_list, function(x) {linear_mdl_tmp(met = x)}) %>% rbindlist()
        res <- merge(res, sph_info, by = "metabolite")
        res[, fdr_bh := p.adjust(pval, method = "BH")] 
        
        setcolorder(res, first_cols)
        res <- res[order(pval)]
        return(res)
}

```

## 3.2. Association with # of ICS prescriptions

```{r}

all_subj_asth <- sph_pheno[SampleGroup == "Asthmatic", ]
all_subj_asth[, .N, .(is.na(bmi), ocs_ard_serum_coll)]

# volcano_plot_fnc(res = linear_res_fnc(expo = "ics_trim_totnum_1y", mets_list = sph_list, 
#                                       dat = "all_subj_asth"), 
#                  title = "Metabolites associations with # of ICS prescriptions w/in 1yr before serum collection", 
#                  subtitle = "adjusted for age, gender, race", 
#                  sig_cutoff = "pval")

volcano_plot_fnc(res = linear_res_fnc(expo = "ics_trim_totnum_1y", mets_list = sph_list, 
                                      dat = "all_subj_asth", add_covar = " + bmi"), 
                 title = "Metabolites associations with # of ICS prescriptions w/in 1yr before serum collection", 
                 subtitle = "adjusted for age, gender, race, BMI closest to serum collection", 
                 sig_cutoff = "pval")

volcano_plot_fnc(res = linear_res_fnc(expo = "ics_trim_totnum_1y", mets_list = sph_list, 
                                      dat = "all_subj_asth[ocs_ard_serum_coll == 0, ]", add_covar = " + bmi"), 
                 title = "Metabolites associations with # of ICS prescriptions w/in 1yr before serum collection", 
                 subtitle = "adjusted for age, gender, race, BMI closest to serum collection; removed OCS w/in 10d", 
                 sig_cutoff = "pval")

```

## 3.3. Association with # of OCS prescriptions

```{r}

# volcano_plot_fnc(res = linear_res_fnc(expo = "ocs_trim_totnum_1y", mets_list = sph_list, 
#                                       dat = "all_subj_asth"), 
#                  title = "Metabolites associations with # of OCS prescriptions w/in 1yr before serum collection", 
#                  subtitle = "adjusted for age, gender, race", 
#                  sig_cutoff = "pval")

volcano_plot_fnc(res = linear_res_fnc(expo = "ocs_trim_totnum_1y", mets_list = sph_list, 
                                      dat = "all_subj_asth", add_covar = " + bmi"), 
                 title = "Metabolites associations with # of OCS prescriptions w/in 1yr before serum collection", 
                 subtitle = "adjusted for age, gender, race, BMI closest to serum collection", 
                 sig_cutoff = "pval")

volcano_plot_fnc(res = linear_res_fnc(expo = "ocs_trim_totnum_1y", mets_list = sph_list, 
                                      dat = "all_subj_asth[ocs_ard_serum_coll == 0, ]", add_covar = " + bmi"), 
                 title = "Metabolites associations with # of OCS prescriptions w/in 1yr before serum collection", 
                 subtitle = "adjusted for age, gender, race, BMI closest to serum collection; removed OCS w/in 10d", 
                 sig_cutoff = "pval")

```

# 3. Save results

```{r, eval = F}

save(list = c(grep("^res_", ls(), value = T), grep("y$", ls(), value = T), "sph_info", "sph_list"), 
     file = here(res_dir, "sph_results.RData"))

```

# 4. Session info

```{r}

sessionInfo()

```
