---
title: "Choosing MGB Biobank asthmatic subject/sample numbers for further omics study"
subtitle: "Demographics, lung function & medication"
author: "MHuang"
date: "2/8/2021"
output:   
  html_document: 
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

## Packages

pkg <- c("tidyverse", "data.table", "here", "DT", "readxl", "grid", "gridExtra")
for (p in pkg) {
        if (require(p, character.only = T)) {
                print(paste0(p, " loaded successfully"))
        } else {
                install.packages(p)
                require(p, character.only = T)
                print(paste0(p, " downloaded and loaded successfully"))
        }
}


## Paths

asthma_dir <- "C:/Users/MHuang/Dropbox (Partners HealthCare)/Asthma/Cleaned_Generated_Files/2020-11-09/"
asthma_rpdr_dir <- "C:/Users/MHuang/Dropbox (Partners HealthCare)/Asthma/RPDR_Downloads/2020-06-30/"
pegasus_dir <- "C:/Users/MHuang/Dropbox (Partners HealthCare)/PEGASUS/rpdr_data/"

dat_dir <- "data/"
fig_dir <- "figures/"
res_dir <- "results/"


## File names

asthma_fname <- str_c(asthma_dir, "RPDR_cleaned_data_2020-11-09.csv")
pegasus1_id_fname <- str_c(pegasus_dir, "PEGASUS_Subject_IDs_n1140.csv")
rpdr_file_header <- str_c(asthma_rpdr_dir, "mns43_063020114847776198_")
rpdr_file_ending <- ".txt"
med_map_fname <- str_c(pegasus_dir, "Medication_Mapping.xlsx")
plasma_dates_fname <- here(dat_dir, "Asthma_Plasma_BBIds_Analysis_Actual_Collection_Dates.xlsx")


## Functions

source(str_c(pegasus_dir, "rpdr_data_wrangling/scripts/schu/fxns.rpdr.drv.phen_1022_MHedit.R"))


## Hard-coded numbers 

sig_digits <- 3

```

# 1. Read in lists and exclude 

Exclude:  

 - existing PEGASUS subjects  
 - subjects with no plasma sample  
 - non-asthmatics  

```{r}

asthma <- fread(asthma_fname); nrow(asthma)
pegasus1_id <- fread(pegasus1_id_fname)

asthma[, ':='(in_pegasus1 = ifelse(Biobank_Subject_ID %in% pegasus1_id$SubjectID, 1, 0), 
              plasma_existence = Plasma_Existence_Yes_No_To_6_30_2020, 
              asthma_existence = Asthma_current_or_past_history_custom_PPV_0_80PPV_Existence_Yes_No, 
              asthma_ppv = Asthma_current_or_past_history_custom_PPV_0_80PPV_List_of_All_Values, 
              copd_existence = COPD_current_or_past_history_custom_PPV_0_75PPV_Existence_Yes_No, 
              copd_ppv = COPD_current_or_past_history_custom_PPV_0_75PPV_List_of_All_Values)]

asthma[, .N, .(in_pegasus1, plasma_existence, asthma_existence)][order(in_pegasus1, plasma_existence)]
asthma[plasma_existence == "Yes" & asthma_existence == "Yes", .N, 
       .(has_first_colldate = !is.na(Plasma_Date_First_To_6_30_2020), 
         has_last_colldate = !is.na(Plasma_Date_Most_Recent_To_6_30_2020), 
         missing_asthma_ppv = is.na(asthma_ppv), asthma_ppv_gt85 = asthma_ppv > 0.85, 
         asthma_ppv_gt90 = asthma_ppv > 0.9, asthma_ppv_gt95 = asthma_ppv > 0.95)]


## Plasma collection dates from Biobank

plasma_dates <- read_xlsx(plasma_dates_fname) %>% 
        as.data.table() %>% 
        setnames(c("SUBJECTID", "Plasma (ml)"), c("Biobank_Subject_ID", "Plasma_ml"))
plasma_dates[, ':='(Biobank_Subject_ID = as.integer(Biobank_Subject_ID), 
                    COLLECT_DATE = as.Date(COLLECT_DATE, format = "%y-%m-%d"))]
setkey(plasma_dates, Biobank_Subject_ID, COLLECT_DATE)

plasma_dates_per_subj <- plasma_dates[, .(n_plasma_date = .N, 
                                          all_plasma_date = str_c(COLLECT_DATE, collapse = ";"), 
                                          first_plasma_date = head(COLLECT_DATE, 1), 
                                          last_plasma_date = tail(COLLECT_DATE, 1)), 
                                      by = Biobank_Subject_ID]

asthma <- merge(asthma, plasma_dates_per_subj, by = "Biobank_Subject_ID", all.x = T)
asthma[, .N, .(plasma_existence, has_plasma_collection_date = !is.na(n_plasma_date))]


## Generate potential subjects selection pool 

tmp_cols <- c(colnames(asthma)[1:which(colnames(asthma) == "Age_Range")], "in_pegasus1", 
              "plasma_existence", "asthma_existence", "asthma_ppv", "copd_existence", "copd_ppv", 
              colnames(plasma_dates_per_subj)[-1])
pool <- asthma[asthma_existence == "Yes" & !is.na(n_plasma_date), 
               ..tmp_cols]; nrow(pool)


## Demographics

pool[, summary(Age)]
ggplot(pool, aes(x = Gender, y = Age)) + 
        geom_jitter(alpha = 0.5) + 
        geom_violin(alpha = 0.5)

pool[, race := case_when(Race %in% c("BLACK OR AFRICAN AMERICAN", "Black-BLACK", 
                                     "Hispanic Black-BLACK OR AFRICAN AMERICAN@HISPANIC") ~ "Black",
                         Race %in% c("White-WHITE", "Hispanic White-WHITE@HISPANIC") ~ "White", 
                         Race %in% c("Other-OTHER") ~ "Other", 
                         Race %in% c("Not Recorded-DECLINED", "Not Recorded-UNKNOWN", "Not Recorded-U", 
                                     "Not Recorded-@", "Not Recorded-NOT GIVEN") ~ "Unknown", 
                         Race %in% c("Hispanic-HISPANIC OR LATINO", "Hispanic-OTHER@HISPANIC", 
                                     "Hispanic-HISPANIC OR LATINO@HISPANIC", "Hispanic-HISPANIC", 
                                     "Hispanic-DECLINED@HISPANIC", "Hispanic-HISPANIC@HISPANIC", 
                                     "Hispanic-@HISPANIC") ~ "Hispanic", 
                         Race %in% c("Asian-ASIAN") ~ "Asian", 
                         Race %in% c("American Indian-AMERICAN INDIAN OR ALASKA NATIVE") ~ "American Indian/Alaska Native", 
                         Race %in% c("Hawaiian-NATIVE HAWAIIAN OR OTHER PACIFIC ISLANDER") ~ "Hawaiian/Pacific Islander")]
pool[, .N, race][, pct := N/sum(N)*100] %>% print(digits = sig_digits)


## COPD existence

pool[, summary(copd_ppv)]
pool[, .N, .(copd_existence, copd_ppv_gt80 = copd_ppv > 0.80, copd_ppv_gt85 = copd_ppv > 0.85, 
             copd_ppv_gt90 = copd_ppv > 0.9)][order(copd_existence, copd_ppv_gt80, copd_ppv_gt85, copd_ppv_gt90)]

```

# 2. Lung function data availability

Variable names:  

 - n_pul_rec: number of pulmonary function test (pft, spirometry) records  
 - earliest_pul_rec: date of earliest pft record  
 - latest_pul_rec: date of latest pft record  
 - has_pul_rec: has any pft record  

Suffix:  

 - _alltime: all time in system  
 - _bflastpl: all time prior to last plasma date  
 - _10ybflastpl: within 10 years prior to last plasma date  
 - _5ybflastpl: within 5 years prior to last plasma date  

## 2.1. All available records

```{r}

filePathToPulTxt <- str_c(rpdr_file_header, "Pul_v2", rpdr_file_ending)
pul_id <- getPULIDs(filePathToPulTxt); nrow(pul_id)
print(str_c("Unique number of subjects with pft data in all biobank asthmatics: ", 
            pul_id[, length(unique(EMPI))]))


## Restrict to those with plasma

pul_id <- pul_id[EMPI %in% pool$EMPI, ]; nrow(pul_id)
print(str_c("Unique number of asthmatics with pft data & plasma sample & not in PEGASUS: ", 
            pul_id[, length(unique(EMPI))]))
pul_id <- merge(pul_id, pool[, .(EMPI, last_plasma_date)], by = "EMPI", all.x = T)


## Calculate timegap between pul record & last plasma date

pul_id[, pul_timegap_d := as.numeric(Date_Fmt - last_plasma_date)][, pul_timegap_y := pul_timegap_d/365.25]
ggplot(pul_id) + 
        geom_histogram(aes(pul_timegap_y), bins = 200) + 
        labs(title = "Time gap in years between PFT record and last plasma date (before = -)")


## Function to generate variables on number of pul records under certain restriction

n_pul_rec_fnc <- function(restriction, varname_suffix) {
        pul_uniq_date <- eval(parse(text = str_c("unique(pul_id[", restriction, 
                                                 ", .(EMPI, Date_Fmt, pul_timegap_d)])")))
        pul_per_subj <- pul_uniq_date[, .(n_pul_rec = .N, 
                                          earliest_pul_rec = min(Date_Fmt), 
                                          latest_pul_rec = max(Date_Fmt)), EMPI]
        pul_per_subj <- merge(pul_per_subj, 
                              pul_uniq_date[pul_timegap_d < 0, 
                                            .(closest_pul_rec_bf_timegap_d = max(pul_timegap_d)), EMPI], 
                              by = "EMPI", all.x = T)
        pul_per_subj <- merge(pul_per_subj, 
                              pul_uniq_date[pul_timegap_d >= 0, 
                                            .(closest_pul_rec_af_timegap_d = min(pul_timegap_d)), EMPI], 
                              by = "EMPI", all.x = T)
        setnames(pul_per_subj, c("EMPI", str_c(colnames(pul_per_subj)[-1], varname_suffix)))
        pool <- merge(pool, pul_per_subj, by = "EMPI", all.x = T)
        eval(parse(text = str_c("pool[, has_pul_rec", varname_suffix, 
                                " := ifelse(!is.na(n_pul_rec", varname_suffix, "), 1, 0) %>% as.factor()]")))
        return(pool)
}


## Generate variables on number of pul records from all time in MGB system

pool <- n_pul_rec_fnc(restriction = "", varname_suffix = "_alltime")


## Examine data availability - all time in MGB system

# display_cols <- c("EMPI", "last_plasma_date", grep("_alltime", colnames(pool), value = T))
# datatable(pool[, ..display_cols], filter = "top")

pool[n_pul_rec_alltime >= 2, .N, .(has_pul_1ybf = closest_pul_rec_bf_timegap_d_alltime >= -365)] 
pool[n_pul_rec_alltime >= 2, .N, .(has_pul_1yaf = closest_pul_rec_af_timegap_d_alltime <= 365)]
pool[n_pul_rec_alltime >= 2, ][closest_pul_rec_bf_timegap_d_alltime >= -365 | 
                                 closest_pul_rec_af_timegap_d_alltime <= 365, .N]
pool[n_pul_rec_alltime >= 2, ][closest_pul_rec_bf_timegap_d_alltime >= -365 & 
                                 closest_pul_rec_af_timegap_d_alltime <= 365, .N]

```

## 2.2. All records before last plasma date

```{r}

pool <- n_pul_rec_fnc(restriction = "pul_timegap_d <= 0", varname_suffix = "_bflastpl")


## Examine data availability - before last plasma collection date

# display_cols <- c("EMPI", "last_plasma_date", grep("_bflastpl", colnames(pool), value = T))
# datatable(pool[, ..display_cols], filter = "top")

```

## 2.3. All records within 10 years before last plasma date

```{r}

pool <- n_pul_rec_fnc(restriction = "pul_timegap_d <= 0 & pul_timegap_y >= -10", 
                      varname_suffix = "_10ybflastpl")


## Examine data availability - within 10 years before last plasma collection date

# display_cols <- c("EMPI", "last_plasma_date", grep("_10ybflastpl", colnames(pool), value = T))
# datatable(pool[, ..display_cols], filter = "top")

```

## 2.4. All records within 5 years before last plasma date

```{r}

pool <- n_pul_rec_fnc(restriction = "pul_timegap_d <= 0 & pul_timegap_y >= -5", 
                      varname_suffix = "_5ybflastpl")


## Examine data availability - within 5 years before last plasma collection date

# display_cols <- c("EMPI", "last_plasma_date", grep("_5ybflastpl", colnames(pool), value = T))
# datatable(pool[, ..display_cols], filter = "top")

```

## 2.5. Compare data availability in different time frames

```{r}

pool[, .N, .(has_pul_rec_alltime, has_pul_rec_bflastpl, has_pul_rec_10ybflastpl, has_pul_rec_5ybflastpl)
     ][order(has_pul_rec_alltime, has_pul_rec_bflastpl, has_pul_rec_10ybflastpl, has_pul_rec_5ybflastpl)]

pool[n_pul_rec_alltime >= 2 & Age >= 18, .N, .(has_pul_1ybf = closest_pul_rec_bf_timegap_d_alltime >= -365)] 
pool[n_pul_rec_alltime >= 2 & Age >= 18, .N, .(has_pul_1yaf = closest_pul_rec_af_timegap_d_alltime <= 365)]
pool[n_pul_rec_alltime >= 2 & Age >= 18, ][closest_pul_rec_bf_timegap_d_alltime >= -365 | 
                                             closest_pul_rec_af_timegap_d_alltime <= 365, .N]
pool[n_pul_rec_alltime >= 2 & Age >= 18, ][closest_pul_rec_bf_timegap_d_alltime >= -365 & 
                                             closest_pul_rec_af_timegap_d_alltime <= 365, .N]

```

# 3. Medication data

Variable names:  

 - n_ics_rec: number of ICS prescription records  
 - ics_rec_latest_timegap_d: date of latest ICS prescription - last plasma date in days
 - has_ics_rec: has ICS prescription records  
 - n_ocs_rec: number of OCS prescription records  
 - ocs_rec_latest_timegap_d: date of latest OCS prescription - last plasma date in days
 - has_ocs_rec: has OCS prescription records  

Suffix:  

 - _1ybflastpl: within 1 year prior to last plasma date  
 - _2ybflastpl: within 2 years prior to last plasma date  
 - _3ybflastpl: within 3 years prior to last plasma date  
 - _4ybflastpl: within 4 years prior to last plasma date  
 - _5ybflastpl: within 5 years prior to last plasma date  
 - _10ybflastpl: within 10 years prior to last plasma date  
 - _bflastpl: all time prior to last plasma date  
 - _alltime: all time in system  

```{r}

med <- fread(str_c(rpdr_file_header, "Med", rpdr_file_ending))
med[, unique(EMPI)] %in% pool[, EMPI] %>% table()
med <- med[EMPI %in% pool$EMPI, ] # Restrict to subjects in selection pool

med[, Medication_Date := as.Date(Medication_Date, format = "%m/%d/%Y")]
med[, summary(Medication_Date)]

med <- merge(med, pool[, .(EMPI, last_plasma_date)], by = "EMPI", all.x = T)
med[, med_timegap_d := as.numeric(Medication_Date - last_plasma_date)][, med_timegap_y := med_timegap_d/365.25]
med[, summary(med_timegap_y)]

```

## 3.1. Parse out corticosteroids

```{r}

## Mapping file for medications of interest

med_map <- read_xlsx(med_map_fname) %>% as.data.table()
med_map_cs <- med_map[Medication_GWAS_Group %in% c("Oral Corticosteroids", "Inhaled Corticosteroids", 
                                                   "Inhaled Corticosteroids and Long Acting Beta Agonists") | 
                        Medication_Pegasus_Group %in% c("Oral Steroids", "Inhaled Steroids", "Cortisone", 
                                                        "Hydrocortisone"), ]
datatable(med_map_cs[, .(Medication_Name, Medication_GWAS_Group, Medication_Pegasus_Group, 
                         ICS_or_ICS_LABA, Medication_Biobank_Folder)])

med_condensed <- med %>% group_by(Medication) %>% summarise()
med_all_subgroups <- data.frame(Medication = character(),
                                Medication_Name = character(),
                                Medication_Biobank_Folder = character(),
                                Medication_GWAS_Group = character(),
                                Medication_Pegasus_Group = character())
for (m in 1:nrow(med_map_cs)) {
        med_subSampleGroup <- med_condensed %>% filter(grepl(med_map_cs[m, ]$Search_Term,
                                                             Medication,
                                                             ignore.case = med_map_cs[m, ]$Ignore.case,
                                                             perl = med_map_cs[m, ]$Perl)) %>%
                mutate(Medication_Name  = med_map_cs[m, ]$Medication_Name,
                       Medication_Biobank_Folder = med_map_cs[m, ]$Medication_Biobank_Folder,
                       Medication_GWAS_Group = med_map_cs[m, ]$Medication_GWAS_Group,
                       Medication_Pegasus_Group = med_map_cs[m, ]$Medication_Pegasus_Group)
        med_all_subgroups <- rbind(med_all_subgroups, med_subSampleGroup)
        rm(med_subSampleGroup)
}
med_all_subgroups <- as.data.table(med_all_subgroups)


## Remove duplicated records

med_all_subgroups[Medication %in% med_all_subgroups[duplicated(Medication), Medication], ]
med_all_subgroups <- med_all_subgroups[!(grepl("Pulmicort flexhaler 180 mcg", Medication) &
                                           Medication_Name == "Dexamethasone")]
med_all_subgroups <- med_all_subgroups[!(grepl("Id-Fluticasone Or Fluticasone/Salmeterol", Medication) &
                                           Medication_Name == "Fluticasone/salmeterol")]


## Add indicator for ICS/OCS/(hydro)cortisone

med_all_subgroups[, ':='(ocs = ifelse(Medication_GWAS_Group == "Oral Corticosteroids" | 
                                        Medication_Pegasus_Group == "Oral Steroids", 1, 0), 
                         ics = ifelse(Medication_GWAS_Group %in% c("Inhaled Corticosteroids", 
                                                                   "Inhaled Corticosteroids and Long Acting Beta Agonists") | 
                                        Medication_Pegasus_Group == "Inhaled Steroids", 1, 0), 
                         cortisone_hydro = ifelse(Medication_Pegasus_Group %in% c("Cortisone", "Hydrocortisone"), 1, 0))]

med_cs <- merge(med, med_all_subgroups, by = "Medication")
med_cs[, unique(EMPI)] %in% pool[, EMPI] %>% table() # Number of subjects with corticosteroids med data
med_cs[med_timegap_y <= 0, length(unique(EMPI))] # Number of subjects with corticosteroids med data before last plasma date

```

## 3.2. Number of ICS/OCS prescriptions prior to last plasma date
### 3.2.1. Within 1 year prior to last plasma date

```{r}

## Function to generate variables on ICS/OCS under certain restriction

cs_vars_fnc <- function(restriction, varname_suffix) {
        cs_tmp <- eval(parse(text = str_c("med_cs[med_timegap_y <= 0 & ", restriction, ", ]")))
        cs_per_subj <- cs_tmp[, .(n_ics_rec = sum(ics, na.rm = T), 
                                  n_ocs_rec = sum(ocs, na.rm = T)), EMPI]
        cs_per_subj <- merge(cs_per_subj, cs_tmp[ics == 1, .(ics_rec_latest_timegap_d = max(med_timegap_d)), EMPI], 
                             by = "EMPI", all.x = T)
        cs_per_subj <- merge(cs_per_subj, cs_tmp[ocs == 1, .(ocs_rec_latest_timegap_d = max(med_timegap_d)), EMPI], 
                             by = "EMPI", all.x = T)
        setnames(cs_per_subj, c("EMPI", str_c(colnames(cs_per_subj)[-1], varname_suffix)))
        pool <- merge(pool, cs_per_subj, by = "EMPI", all.x = T)
        eval(parse(text = str_c("pool[, ':='(has_ics_rec", varname_suffix, " = ifelse(!is.na(n_ics_rec", 
                                             varname_suffix, "), 1, 0) %>% as.factor(), 
                                             has_ocs_rec", varname_suffix, " = ifelse(!is.na(n_ocs_rec", 
                                             varname_suffix, "), 1, 0) %>% as.factor())]")))
        return(pool)
}


## Generate # of ICS/OCS prescriptions within 1 year prior to last plasma date

pool <- cs_vars_fnc(restriction = "med_timegap_y >= -1", varname_suffix = "_1ybflastpl")

```

### 3.2.2. Within 2 years prior to last plasma date

```{r}

## Generate # of ICS/OCS prescriptions within 2 years prior to last plasma date

pool <- cs_vars_fnc(restriction = "med_timegap_y >= -2", varname_suffix = "_2ybflastpl")

```

### 3.2.3. Within 3 years prior to last plasma date

```{r}

## Generate # of ICS/OCS prescriptions within 3 years prior to last plasma date

pool <- cs_vars_fnc(restriction = "med_timegap_y >= -3", varname_suffix = "_3ybflastpl")

```

### 3.2.4. Within 4 years prior to last plasma date

```{r}

## Generate # of ICS/OCS prescriptions within 4 years prior to last plasma date

pool <- cs_vars_fnc(restriction = "med_timegap_y >= -4", varname_suffix = "_4ybflastpl")

```

### 3.2.5. Within 5 years prior to last plasma date

```{r}

## Generate # of ICS/OCS prescriptions within 5 years prior to last plasma date

pool <- cs_vars_fnc(restriction = "med_timegap_y >= -5", varname_suffix = "_5ybflastpl")

```

# 4. Save selection pool data

```{r}

fwrite(pool, here(res_dir, "selection_pool.csv"))

```

# 5. Session info

```{r}

sessionInfo()

```
