---
title: "Choosing MGB Biobank asthmatic subject/samples for further omics study"
subtitle: "Sample size by varying criteria"
author: "MHuang"
date: "2/18/2021"
output:   
  html_document: 
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

## Packages

pkg <- c("tidyverse", "data.table", "here", "DT", "grid", "gridExtra")
for (p in pkg) {
        if (require(p, character.only = T)) {
                print(paste0(p, " loaded successfully"))
        } else {
                install.packages(p)
                require(p, character.only = T)
                print(paste0(p, " downloaded and loaded successfully"))
        }
}


## Paths

dat_dir <- "data/"
fig_dir <- "figures/"
res_dir <- "results/"


## File names

pool_fname <- here(res_dir, "selection_pool.csv")


## Hard-coded numbers 

sig_digits <- 3

```

# 1. Read in selection pool

Exclude:  

 - subjects with no plasma sample  
 - non-asthmatics  

```{r}

pool <- fread(pool_fname); nrow(pool)

```

Lung function data variable names:  

 - n_pul_rec: number of pulmonary function test (pft, spirometry) records  
 - earliest_pul_rec: date of earliest pft record  
 - latest_pul_rec: date of latest pft record  
 - has_pul_rec: has any pft record  

Medication data variable names:  

 - n_ics_rec: number of ICS prescription records  
 - ics_rec_latest_timegap_d: date of latest ICS prescription - last plasma date in days
 - has_ics_rec: has ICS prescription records  
 - n_ocs_rec: number of OCS prescription records  
 - ocs_rec_latest_timegap_d: date of latest OCS prescription - last plasma date in days
 - has_ocs_rec: has OCS prescription records  

Asthma exacerbations & severity, adrenal insufficiency variable names:  

 - n_exac_dia: number of asthma exacerbation diagnoses    
 - has_exac_dia: has asthma exacerbation diagnoses    
 - n_sevr_dia: number of asthma severity diagnoses  
 - has_sevr_dia: has asthma severity diagnoses  
 - n_adrinsuf_dia: number of adrenal insufficiency diagnoses  
 - has_adrinsuf_dia: has adrenal insufficiency diagnoses  

Suffix:  

 - _alltime: all time in system  
 - _bflastpl: all time prior to last plasma date  
 - _10ybflastpl: within 10 years prior to last plasma date  
 - _5ybflastpl: within 5 years prior to last plasma date  
 - _4ybflastpl: within 4 years prior to last plasma date  
 - _3ybflastpl: within 3 years prior to last plasma date  
 - _2ybflastpl: within 2 years prior to last plasma date  
 - _1ybflastpl: within 1 year prior to last plasma date  

BMI/lab test data variable names (bmi as example, replace with acth/cortisol/eos/ige):  

 - bmi: BMI value of a single record  
 - bmi_date: date of BMI record  
 - n_bmi_rec: number of BMI records 
 - bmi_all_values: list of all BMI values  
 - bmi_all_dates: list of all BMI record dates  
 - bmi_median: median BMI among all values  
 - bmi_min: minimum BMI among all values  
 - bmi_max: maximum BMI among all values  
 - bmi_mean: mean BMI averaged across all values  
 - bmi_closest_timegap_d: date of closest BMI record - last plasma date in days

Suffix:  

 - _closest_lastpl: closest record to last plasma date  
 - _pm3y_lastpl: plus or minus 3 years around last plasma date  
 - _pm2y_lastpl: plus or minus 2 years around last plasma date  
 - _pm1y_lastpl: plus or minus 1 year around last plasma date  

Others:  

 - Age  
 - Gender  
 - race  
 - asthma_ppv: range 0.82 - 0.95  
 - copd_existence: Yes/No (Biobank algorithm)  
 - copd_ppv: range 0.75 - 0.94
 - smk_ever_alltime: ever has a record of former/current smoking during all time in MGB system

# 2. General demographics
## 2.1. All subjects in pool

```{r}

pool[, summary(Age)]
pool[, .N, .(`age>=18` = Age >= 18)]

ggplot(pool, aes(x = Gender, y = Age)) + 
        geom_jitter(alpha = 0.5) + 
        geom_violin(alpha = 0.5)

pool[, .N, race][order(-N)][, pct := N/sum(N)*100] %>% print(digits = sig_digits - 1)

pool[, .N, copd_existence]

pool[, .N, smk_ever_alltime]


## Median BMI from within (+/-)1 year against BMI closest to last plasma date

pool[, .N, .(has_bmi_closest_lastpl, has_bmi_rec_pm1y_lastpl)]
pool[, summary(bmi_closest_timegap_d_pm1y_lastpl)]
ggplot(pool) + geom_histogram(aes(x = bmi_closest_timegap_d_pm1y_lastpl), bins = 200)
ggplot(pool, aes(x = bmi_closest_lastpl, y = bmi_median_pm1y_lastpl)) + 
        geom_point(alpha = 0.5)

```

# 3. Pulmanory function test (pft) data availability
## 3.1. 2+ pft records

```{r}

pool[Age >= 18 & n_pul_rec_alltime >= 2, .N]
pool[Age >= 18 & n_pul_rec_alltime >= 2, summary(Age)]

ggplot(pool[Age >= 18 & n_pul_rec_alltime >= 2, ], aes(x = Gender, y = Age)) + 
        geom_jitter(alpha = 0.5) + 
        geom_violin(alpha = 0.5)

pool[Age >= 18 & n_pul_rec_alltime >= 2, .N, race
     ][order(-N)][, pct := N/sum(N)*100] %>% print(digits = sig_digits - 1)

pool[Age >= 18 & n_pul_rec_alltime >= 2, .N, copd_existence]

pool[Age >= 18 & n_pul_rec_alltime >= 2, .N, smk_ever_alltime]


## Median BMI from within (+/-)1 year against BMI closest to last plasma date

pool[Age >= 18 & n_pul_rec_alltime >= 2, .N, .(has_bmi_closest_lastpl, has_bmi_rec_pm1y_lastpl)]
pool[Age >= 18 & n_pul_rec_alltime >= 2, summary(bmi_closest_timegap_d_pm1y_lastpl)]
ggplot(pool[Age >= 18 & n_pul_rec_alltime >= 2, ]) + 
        geom_histogram(aes(x = bmi_closest_timegap_d_pm1y_lastpl), bins = 200)
ggplot(pool[Age >= 18 & n_pul_rec_alltime >= 2, ], 
       aes(x = bmi_closest_lastpl, y = bmi_median_pm1y_lastpl)) + 
        geom_point(alpha = 0.5)


## Timing of pul record

pool[n_pul_rec_alltime >= 2 & Age >= 18, .N, .(has_pul_1ybf = closest_pul_rec_bf_timegap_d_alltime >= -365)] 
pool[n_pul_rec_alltime >= 2 & Age >= 18, .N, .(has_pul_1yaf = closest_pul_rec_af_timegap_d_alltime <= 365)]
pool[n_pul_rec_alltime >= 2 & Age >= 18, ][closest_pul_rec_bf_timegap_d_alltime >= -365*5 | 
                                             closest_pul_rec_af_timegap_d_alltime <= 365*5, .N]
pool[n_pul_rec_alltime >= 2 & Age >= 18, ][closest_pul_rec_bf_timegap_d_alltime >= -365 & 
                                             closest_pul_rec_af_timegap_d_alltime <= 365, .N]


## Timegap between 1st and last pul records

pool[n_pul_rec_alltime >= 2 & Age >= 18, 
     timegap_y_1st_last_pul := as.numeric(as.Date(earliest_pul_rec_alltime, "%Y-%m-%d") - 
                                            as.Date(latest_pul_rec_alltime, "%Y-%m-%d"))/365.25]
ggplot(pool[n_pul_rec_alltime >= 2 & Age >= 18, ]) + 
        geom_histogram(aes(timegap_y_1st_last_pul), bins = 200)
pool[n_pul_rec_alltime >= 2 & Age >= 18, .N, .(timegap_y_1st_last_pul <= -1)]

pool[Age >= 18 & n_pul_rec_alltime >= 2, .(in_pegasus1, n_pul_rec_alltime, closest_pul_rec_bf_timegap_d_alltime, 
                                           closest_pul_rec_af_timegap_d_alltime, timegap_y_1st_last_pul, 
                                           has_bmi_closest_lastpl, has_bmi_rec_pm1y_lastpl, 
                                           asthma_ppv, Age, copd_existence, smk_ever_alltime, 
                                           n_ics_rec_5ybflastpl, n_ocs_rec_5ybflastpl, n_exac_dia_5ybflastpl, 
                                           n_sevr_dia_5ybflastpl, n_ige_rec_pm1y_lastpl, n_cortisol_rec_pm1y_lastpl, 
                                           n_acth_rec_pm1y_lastpl, n_eos_rec_pm1y_lastpl)] %>% 
        datatable(filter = "top")

```

## 3.2. 1 pft record

```{r}

pool[Age >= 18 & n_pul_rec_alltime == 1, .N]
pool[Age >= 18 & n_pul_rec_alltime == 1, summary(Age)]

ggplot(pool[Age >= 18 & n_pul_rec_alltime == 1, ], aes(x = Gender, y = Age)) + 
        geom_jitter(alpha = 0.5) + 
        geom_violin(alpha = 0.5)

pool[Age >= 18 & n_pul_rec_alltime == 1, .N, race
     ][order(-N)][, pct := N/sum(N)*100] %>% print(digits = sig_digits - 1)

pool[Age >= 18 & n_pul_rec_alltime == 1, .N, copd_existence]

pool[Age >= 18 & n_pul_rec_alltime == 1, .N, smk_ever_alltime]


## Median BMI from within (+/-)1 year against BMI closest to last plasma date

pool[Age >= 18 & n_pul_rec_alltime == 1, .N, .(has_bmi_closest_lastpl, has_bmi_rec_pm1y_lastpl)]
pool[Age >= 18 & n_pul_rec_alltime == 1, summary(bmi_closest_timegap_d_pm1y_lastpl)]
ggplot(pool[Age >= 18 & n_pul_rec_alltime == 1, ]) + 
        geom_histogram(aes(x = bmi_closest_timegap_d_pm1y_lastpl), bins = 200)
ggplot(pool[Age >= 18 & n_pul_rec_alltime == 1, ], 
       aes(x = bmi_closest_lastpl, y = bmi_median_pm1y_lastpl)) + 
        geom_point(alpha = 0.5)


## Timing of pul record

pool[Age >= 18 & n_pul_rec_alltime == 1, .N, .(has_pul_1ybf = closest_pul_rec_bf_timegap_d_alltime >= -365)] 
pool[Age >= 18 & n_pul_rec_alltime == 1, .N, .(has_pul_1yaf = closest_pul_rec_af_timegap_d_alltime <= 365)]
pool[Age >= 18 & n_pul_rec_alltime == 1, ][closest_pul_rec_bf_timegap_d_alltime >= -365 | 
                                             closest_pul_rec_af_timegap_d_alltime <= 365, .N]

pool[Age >= 18 & n_pul_rec_alltime == 1, .(in_pegasus1, n_pul_rec_alltime, closest_pul_rec_bf_timegap_d_alltime, 
                                           closest_pul_rec_af_timegap_d_alltime, timegap_y_1st_last_pul, 
                                           has_bmi_closest_lastpl, has_bmi_rec_pm1y_lastpl, 
                                           asthma_ppv, Age, copd_existence, smk_ever_alltime, 
                                           n_ics_rec_5ybflastpl, n_ocs_rec_5ybflastpl, n_exac_dia_5ybflastpl, 
                                           n_sevr_dia_5ybflastpl, n_ige_rec_pm1y_lastpl, n_cortisol_rec_pm1y_lastpl, 
                                           n_acth_rec_pm1y_lastpl, n_eos_rec_pm1y_lastpl)] %>% 
        datatable(filter = "top")

```

# 4. ICS medication

```{r}

pool[Age >= 18, .N, .(has_ics_rec_5ybflastpl, has_ics_rec_4ybflastpl, has_ics_rec_3ybflastpl, 
                      has_ics_rec_2ybflastpl, has_ics_rec_1ybflastpl)
     ][order(has_ics_rec_5ybflastpl, has_ics_rec_4ybflastpl, has_ics_rec_3ybflastpl, 
             has_ics_rec_2ybflastpl, has_ics_rec_1ybflastpl)]

pool[Age >= 18, .(has_ics_rec_5ybflastpl, has_ics_rec_4ybflastpl, has_ics_rec_3ybflastpl, has_ics_rec_2ybflastpl, 
                  has_ics_rec_1ybflastpl, has_acth_rec_pm1y_lastpl, has_cortisol_rec_pm1y_lastpl, 
                  has_exac_dia_5ybflastpl, n_exac_dia_5ybflastpl, has_sevr_dia_5ybflastpl, n_sevr_dia_5ybflastpl, 
                  has_adrinsuf_dia_5ybflastpl, n_adrinsuf_dia_5ybflastpl, has_pul_rec_5ybflastpl, n_pul_rec_5ybflastpl, 
                  asthma_ppv, Age, copd_existence, has_bmi_closest_lastpl, smk_ever_alltime)] %>% 
        datatable(filter = "top")

```

# 5. Session info

```{r}

sessionInfo()

```
