---
title: "Excerbation variables"
subtitle: "For Priya"
author: "MHuang"
date: "3/20/2021"
output:   
  html_document: 
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

## Packages

pkg <- c("tidyverse", "data.table", "here", "DT", "grid", "gridExtra", "readxl")
for (p in pkg) {
        if (require(p, character.only = T)) {
                print(paste0(p, " loaded successfully"))
        } else {
                install.packages(p)
                require(p, character.only = T)
                print(paste0(p, " downloaded and loaded successfully"))
        }
}


## Paths

asthma_rpdr_dir <- "C:/Users/MHuang/Dropbox (Partners HealthCare)/Asthma/RPDR_Downloads/2020-06-30/"
pegasus_dir <- "C:/Users/MHuang/Dropbox (Partners HealthCare)/PEGASUS/rpdr_data/"

dat_dir <- "data/"
fig_dir <- "figures/"
res_dir <- "results/"


## File names

pool_fname <- here(res_dir, "selection_pool.csv")
rpdr_file_header <- str_c(asthma_rpdr_dir, "mns43_063020114847776198_")
rpdr_file_ending <- ".txt"
med_map_fname <- str_c(pegasus_dir, "Medication_Mapping.xlsx")


## Hard-coded numbers 

sig_digits <- 3

```

# 1. Exacerbations post last plasma date
## 1.1. Subject selection pool & diagnosis data

```{r}

pool <- fread(pool_fname); nrow(pool)
pool[, ':='(first_plasma_date = as.Date(first_plasma_date, format = "%Y-%m-%d"), 
            last_plasma_date = as.Date(last_plasma_date, format = "%Y-%m-%d"))]

dia <- fread(str_c(rpdr_file_header, "Dia", rpdr_file_ending), quote = "")
dia[, unique(EMPI)] %in% pool[, EMPI] %>% table()
dia <- dia[EMPI %in% pool$EMPI, ] # Restrict to subjects in selection pool

dia[, Date := as.Date(Date, format = "%m/%d/%Y")]
dia[, summary(Date)]

dia <- merge(dia, pool[, .(EMPI, first_plasma_date, last_plasma_date)], by = "EMPI", all.x = T)
dia[, ':='(dia_timegap_to1st_d = as.numeric(Date - first_plasma_date), 
           dia_timegap_tolast_d = as.numeric(Date - last_plasma_date))]
dia[, ':='(dia_timegap_to1st_y = dia_timegap_to1st_d/365.25, 
           dia_timegap_tolast_y = dia_timegap_tolast_d/365.25)]
dia[, summary(dia_timegap_to1st_y)]; dia[, summary(dia_timegap_tolast_y)]
dia[, .N, dia_timegap_to1st_d == dia_timegap_tolast_d]

dia_condensed <- unique(dia[, .(Diagnosis_Name, Code_Type, Code)])

```

## 1.2. Asthma exacerbations data

> Status asthmaticus is a medical emergency, an extreme form of asthma exacerbation characterized by hypoxemia, hypercarbia, and secondary respiratory failure (Ref: [NCBI Book](https://www.ncbi.nlm.nih.gov/books/NBK526070/)). 

```{r}

dia_condensed[grep("asthma", Diagnosis_Name, ignore.case = T), 
              ][grep("exacer|asthmaticus", Diagnosis_Name, ignore.case = T), ] %>% datatable()
dia_asth_exac_icd10 <- dia_condensed[grep("asthma", Diagnosis_Name, ignore.case = T), 
                                     ][grep("exacer|asthmaticus", Diagnosis_Name, ignore.case = T), 
                                       ][Code_Type == "ICD10", ]
dia_asth_exac_icd10[, Diagnosis_Name := gsub("[(),]", "", 
                                             gsub(" ", "_", Diagnosis_Name))]


## There may be multiple records for the same person on the same day

asth_exac <- dia[grep("asthma", Diagnosis_Name, ignore.case = T), 
                 ][grep("exacer|with status asthmaticus", Diagnosis_Name, ignore.case = T), ]
asth_exac <- unique(asth_exac[, .(EMPI, Date, Diagnosis_Name, Code_Type, Code, 
                                  last_plasma_date, dia_timegap_tolast_d, dia_timegap_tolast_y)])
asth_exac[, Diagnosis_Name := gsub("[(),]", "", 
                                   gsub(" ", "_", Diagnosis_Name))]

```

## 1.3. Generate relevant variables on or post last plasma date

```{r}

dia_asth_exac_icd10$Diagnosis_Name
exac_tmp <- asth_exac[dia_timegap_tolast_y >= 0, ]

exac_vars_fnc <- function() {
        
        dat <- pool[, .(EMPI, Biobank_Subject_ID, asthma_existence, asthma_ppv, first_plasma_date, last_plasma_date)]

        # This restrict data to those on or post last plasma date

        for (e in dia_asth_exac_icd10$Diagnosis_Name) {
                
                tmp_per_subj <- exac_tmp[Diagnosis_Name == e, ][, .(event_counts = .N, 
                                                                    event_dates = str_c(Date, collapse = ";"), 
                                                                    event_1st_date = min(Date)), 
                                                                EMPI]
                dat <- merge(dat, tmp_per_subj, by = "EMPI", all.x = T)
                dat[, ':='(Any_event = ifelse(!is.na(event_counts), 1, 0), 
                           Days_to_1st_event = as.numeric(event_1st_date - last_plasma_date))]
                
                setnames(dat, gsub("event", e, colnames(dat)))
                setnames(dat, grep(e, colnames(dat), value = T), 
                         str_c("Dia_PostDate_", grep(e, colnames(dat), value = T)))
        }

        return(dat)
}

exac_dat <- exac_vars_fnc()

fwrite(exac_dat, file = here(res_dir, "exacerbation_data_post_last_plasma_date_all_subjects.csv"))

```

Column names example:

 - Dia_PostDate_Unspecified_asthma_with_acute_exacerbation_counts: number of exacerbations recorded on or post last plasma data
 - Dia_PostDate_Unspecified_asthma_with_acute_exacerbation_dates: dates of exacerbations recorded on or post last plasma data
 - Dia_PostDate_Unspecified_asthma_with_acute_exacerbation_1st_date: date of first exacerbation recorded on or post last plasma data
 - Dia_PostDate_Any_Unspecified_asthma_with_acute_exacerbation: existence of this type pf exacerbation (1=yes; 0=no) on or post last plasma data
 - Dia_PostDate_Days_to_1st_Unspecified_asthma_with_acute_exacerbation: number of days from last plasma date to date of first exacerbation


# 2. Time in RPDR up to first plasma date

Based on diagnosis since we don't have encounters data

```{r}

setkey(dia, EMPI, dia_timegap_to1st_d)
dia_1st_rec <- dia[, head(.SD, 1), EMPI]
dia_1st_rec[, ':='(days_in_rpdr_dia = abs(dia_timegap_to1st_d), 
                   years_in_rpdr_dia = abs(dia_timegap_to1st_y))]

```

# 3. Lab test data before first plasma date

```{r}

lab <- fread(str_c(rpdr_file_header, "Lab", rpdr_file_ending), quote = "")
lab[, unique(EMPI)] %in% pool[, EMPI] %>% table()
lab <- lab[EMPI %in% pool$EMPI, ] # Restrict to subjects in selection pool

lab[, Lab_Date := as.Date(substr(Seq_Date_Time, 1, 10), format = "%m/%d/%Y")]
lab[grepl(":", Seq_Date_Time), ':='(Lab_Hour = gsub("(.*) (.*):(.*)", "\\2", Seq_Date_Time), 
                                    Lab_Minute = gsub("(.*) (.*):(.*)", "\\3", Seq_Date_Time))]
lab[, summary(Lab_Date)]

lab <- merge(lab, pool[, .(EMPI, first_plasma_date, last_plasma_date)], by = "EMPI", all.x = T)
lab[, ':='(lab_timegap_to1st_d = as.numeric(Lab_Date - first_plasma_date), 
           lab_timegap_tolast_d = as.numeric(Lab_Date - last_plasma_date))]
lab[, ':='(lab_timegap_to1st_y = lab_timegap_to1st_d/365.25, 
           lab_timegap_tolast_y = lab_timegap_tolast_d/365.25)]
lab[, summary(lab_timegap_to1st_y)]; lab[, summary(lab_timegap_tolast_y)]
lab[, .N, lab_timegap_to1st_d == lab_timegap_tolast_d]

lab_condensed <- unique(lab[, .(Group_Id, Loinc_Code, Test_Id, Test_Description, Specimen_Type)])

```

## 3.1. Eosinophil 

```{r}

lab_condensed[grepl("(?i)eos", Group_Id), ] %>% datatable()
lab_condensed[grepl("(?i)eos", Group_Id), ][, .N, .(Group_Id, Loinc_Code)]
eos_code <- unique(lab_condensed[grepl("(?i)eos", Group_Id), Loinc_Code])

eos_long <- lab[Loinc_Code %in% eos_code, ]
eos_long[, .N, Reference_Units][order(-N)]
eos_long[, eos := case_when(Reference_Units %in% c("K/uL", "K3/UL", "th/cmm", "th/mm3", "K/UL", "K/MM3", "") ~ as.numeric(Result), 
                            Reference_Units %in% c("/cumm", "/(MM3)", "/CUMM", "UL", "/uL") ~ as.numeric(Result)*0.001)
         ][, eos_unit := "K/uL"] # Unify to K/uL
eos_long[is.na(eos), .N, Result] %>% datatable()
eos_long[, summary(eos)] 

eos_long <- eos_long[!is.na(eos), ]
eos_long[, length(unique(EMPI))]


## Function to generate variables on eos under certain restriction

eos_vars_fnc <- function(restriction, varname_suffix) {
        eos_tmp <- eval(parse(text = str_c("eos_long[", restriction, ", ]")))
        eos_per_subj <- eos_tmp[, .(n_eos_rec = .N, 
                                    eos_all_values = str_c(eos, collapse = ";"), 
                                    eos_all_dates = str_c(Lab_Date, collapse = ";"), 
                                    eos_median = median(eos), 
                                    eos_min = min(eos), 
                                    eos_max = max(eos), 
                                    eos_mean = mean(eos)), EMPI]
        
        setkey(eos_tmp, EMPI, lab_timegap_to1st_d)
        eos_closest <- eos_tmp[, tail(.SD, 1), EMPI]
        setnames(eos_closest, c("eos", "lab_timegap_to1st_d"), 
                              c("eos_closest", "eos_closest_days_to1st"))
        
        eos_per_subj <- merge(eos_per_subj, 
                              eos_closest[, .(EMPI, eos_closest, eos_closest_days_to1st, eos_unit)], 
                              by = "EMPI")
        setnames(eos_per_subj, c("EMPI", str_c(colnames(eos_per_subj)[-1], varname_suffix)))
        return(eos_per_subj)
}


## Generate variables on eos records from within 5/2/1 years before first plasma date

dat <- pool[, .(EMPI, Biobank_Subject_ID, asthma_existence, asthma_ppv, first_plasma_date, last_plasma_date)]
eos_ige_dat <- merge(dat, 
                     eos_vars_fnc(restriction = "lab_timegap_to1st_d <= 0 & lab_timegap_to1st_d >= -5", 
                     varname_suffix = "_5ybf_1stplasma"), 
                     by = "EMPI", all.x = T)
eos_ige_dat <- merge(eos_ige_dat, 
                     eos_vars_fnc(restriction = "lab_timegap_to1st_d <= 0 & lab_timegap_to1st_d >= -2", 
                     varname_suffix = "_2ybf_1stplasma"), 
                     by = "EMPI", all.x = T)
eos_ige_dat <- merge(eos_ige_dat, 
                     eos_vars_fnc(restriction = "lab_timegap_to1st_d <= 0 & lab_timegap_to1st_d >= -1", 
                     varname_suffix = "_1ybf_1stplasma"), 
                     by = "EMPI", all.x = T)

```

## 3.2. IgE 

```{r}

lab_condensed[grepl("(?i)ige", Group_Id), ] %>% datatable()
lab_condensed[grepl("(?i)ige", Group_Id), ][, .N, .(Group_Id, Loinc_Code)]
ige_code <- unique(lab_condensed[grepl("(?i)ige", Group_Id), Loinc_Code])

ige_long <- lab[Loinc_Code %in% ige_code, ]
ige_long[, .N, Reference_Units][order(-N)]
ige_long[, ige := case_when(Reference_Units %in% c("IU/ml", "IU/mL", "kU/L", "IU/ML", "U/ML", "", "mg/dl") ~ as.numeric(Result), 
                            # those with mg/dl as unit has the same reference range as IU/ml (0-100)
                            Reference_Units %in% c("ng/ml") ~ as.numeric(Result)/2.4) # Ref: PMC3206235
         ][, ige_unit := "kU/L"] # Unify to kU/L
ige_long[is.na(ige), .N, Result] %>% datatable()
ige_long[, summary(ige)] 

ige_long <- ige_long[!is.na(ige), ]
ige_long[, length(unique(EMPI))]


## Function to generate variables on eos under certain restriction

ige_vars_fnc <- function(restriction, varname_suffix) {
        ige_tmp <- eval(parse(text = str_c("ige_long[", restriction, ", ]")))
        ige_per_subj <- ige_tmp[, .(n_ige_rec = .N, 
                                    ige_all_values = str_c(ige, collapse = ";"), 
                                    ige_all_dates = str_c(Lab_Date, collapse = ";"), 
                                    ige_median = median(ige), 
                                    ige_min = min(ige), 
                                    ige_max = max(ige), 
                                    ige_mean = mean(ige)), EMPI]
        
        setkey(ige_tmp, EMPI, lab_timegap_to1st_d)
        ige_closest <- ige_tmp[, tail(.SD, 1), EMPI]
        setnames(ige_closest, c("ige", "lab_timegap_to1st_d"), 
                              c("ige_closest", "ige_closest_days_to1st"))
        
        ige_per_subj <- merge(ige_per_subj, 
                              ige_closest[, .(EMPI, ige_closest, ige_closest_days_to1st, ige_unit)], 
                              by = "EMPI")
        setnames(ige_per_subj, c("EMPI", str_c(colnames(ige_per_subj)[-1], varname_suffix)))
        return(ige_per_subj)
}


## Generate variables on ige records from within 5/2/1 years before first plasma date

eos_ige_dat <- merge(eos_ige_dat, 
                     ige_vars_fnc(restriction = "lab_timegap_to1st_d <= 0 & lab_timegap_to1st_d >= -5", 
                     varname_suffix = "_5ybf_1stplasma"), 
                     by = "EMPI", all.x = T)
eos_ige_dat <- merge(eos_ige_dat, 
                     ige_vars_fnc(restriction = "lab_timegap_to1st_d <= 0 & lab_timegap_to1st_d >= -2", 
                     varname_suffix = "_2ybf_1stplasma"), 
                     by = "EMPI", all.x = T)
eos_ige_dat <- merge(eos_ige_dat, 
                     ige_vars_fnc(restriction = "lab_timegap_to1st_d <= 0 & lab_timegap_to1st_d >= -1", 
                     varname_suffix = "_1ybf_1stplasma"), 
                     by = "EMPI", all.x = T)

```

## 3.3. Save data 

```{r}

eos_ige_dat <- merge(eos_ige_dat, 
                     dia_1st_rec[, .(EMPI, days_in_rpdr_dia, years_in_rpdr_dia)], 
                     by = "EMPI", all.x = T)
fwrite(eos_ige_dat, file = here(res_dir, "eos_ige_data_before_first_plasma_date_all_subjects.csv"))

```

Column names (same for eos/ige) in eos_ige_dat: 

 - n_eos_rec: number of eosinophil records within specified time frame
 - eos_all_values: all available eosinophil counts within specified time frame
 - eos_all_dates: all dates of eosinophil records within specified time frame
 - eos_median: median of all eosinophil counts within specified time frame
 - eos_min: minimum of all eosinophil counts within specified time frame
 - eos_max: maximum of all eosinophil counts within specified time frame
 - eos_mean: mean of all eosinophil counts within specified time frame
 - eos_closest: eosinophil count closest to first plasma date within specified time frame
 - eos_closest_days_to1st: number of days between closest eosinophil record and first plasma date within specified time frame
 - eos_unit: unit of eosinophil count
 - days_in_rpdr_dia: days from first diagnosis record to first plasma collection date
 - years_in_rpdr_dia: days_in_rpdr_dia/365.25

Suffix (to specify time frame):

 - _5ybf_1stplasma: five years before first plasma collection date
 - _2ybf_1stplasma: two years before first plasma collection date
 - _1ybf_1stplasma: one year before first plasma collection date

# 4. Medication data

```{r}

med <- fread(str_c(rpdr_file_header, "Med", rpdr_file_ending))
med[, unique(EMPI)] %in% pool[, EMPI] %>% table()
med <- med[EMPI %in% pool$EMPI, ] # Restrict to subjects in selection pool

med[, Medication_Date := as.Date(Medication_Date, format = "%m/%d/%Y")]
med[, summary(Medication_Date)]

med <- merge(med, pool[, .(EMPI, first_plasma_date, last_plasma_date)], by = "EMPI", all.x = T)
med[, ':='(med_timegap_to1st_d = as.numeric(Medication_Date - first_plasma_date), 
           med_timegap_tolast_d = as.numeric(Medication_Date - last_plasma_date))]
med[, ':='(med_timegap_to1st_y = med_timegap_to1st_d/365.25, 
           med_timegap_tolast_y = med_timegap_tolast_d/365.25)]
med[, summary(med_timegap_to1st_y)]; med[, summary(med_timegap_tolast_y)]
med[, .N, med_timegap_to1st_d == med_timegap_tolast_d]

```

## 4.1. Parse out corticosteroids

```{r}

## Mapping file for medications of interest

med_map <- read_xlsx(med_map_fname) %>% as.data.table()
med_map_cs <- med_map[Medication_GWAS_Group %in% c("Oral Corticosteroids", "Inhaled Corticosteroids", 
                                                   "Inhaled Corticosteroids and Long Acting Beta Agonists") | 
                        Medication_Pegasus_Group %in% c("Oral Steroids", "Inhaled Steroids", "Cortisone", 
                                                        "Hydrocortisone"), ]
datatable(med_map_cs[, .(Medication_Name, Medication_GWAS_Group, Medication_Pegasus_Group, 
                         ICS_or_ICS_LABA, Medication_Biobank_Folder)])

med_condensed <- med %>% group_by(Medication) %>% summarise()
med_all_subgroups <- data.frame(Medication = character(),
                                Medication_Name = character(),
                                Medication_Biobank_Folder = character(),
                                Medication_GWAS_Group = character(),
                                Medication_Pegasus_Group = character())
for (m in 1:nrow(med_map_cs)) {
        med_subSampleGroup <- med_condensed %>% filter(grepl(med_map_cs[m, ]$Search_Term,
                                                             Medication,
                                                             ignore.case = med_map_cs[m, ]$Ignore.case,
                                                             perl = med_map_cs[m, ]$Perl)) %>%
                mutate(Medication_Name  = med_map_cs[m, ]$Medication_Name,
                       Medication_Biobank_Folder = med_map_cs[m, ]$Medication_Biobank_Folder,
                       Medication_GWAS_Group = med_map_cs[m, ]$Medication_GWAS_Group,
                       Medication_Pegasus_Group = med_map_cs[m, ]$Medication_Pegasus_Group)
        med_all_subgroups <- rbind(med_all_subgroups, med_subSampleGroup)
        rm(med_subSampleGroup)
}
med_all_subgroups <- as.data.table(med_all_subgroups)


## Remove duplicated records

med_all_subgroups[Medication %in% med_all_subgroups[duplicated(Medication), Medication], ]
med_all_subgroups <- med_all_subgroups[!(grepl("Pulmicort flexhaler 180 mcg", Medication) &
                                           Medication_Name == "Dexamethasone")]
med_all_subgroups <- med_all_subgroups[!(grepl("Id-Fluticasone Or Fluticasone/Salmeterol", Medication) &
                                           Medication_Name == "Fluticasone/salmeterol")]


## Add indicator for ICS/OCS/(hydro)cortisone

med_all_subgroups[, ':='(ocs = ifelse(Medication_GWAS_Group == "Oral Corticosteroids" | 
                                        Medication_Pegasus_Group == "Oral Steroids", 1, 0), 
                         ics = ifelse(Medication_GWAS_Group %in% c("Inhaled Corticosteroids", 
                                                                   "Inhaled Corticosteroids and Long Acting Beta Agonists") | 
                                        Medication_Pegasus_Group == "Inhaled Steroids", 1, 0), 
                         cortisone_hydro = ifelse(Medication_Pegasus_Group %in% c("Cortisone", "Hydrocortisone"), 1, 0))]

med_cs <- merge(med, med_all_subgroups, by = "Medication")
med_cs[, unique(EMPI)] %in% pool[, EMPI] %>% table() # Number of subjects with corticosteroids med data
med_cs[med_timegap_to1st_d <= 0, length(unique(EMPI))] # Number of subjects with corticosteroids med data before first plasma date

```

## 4.2. Number of ICS/OCS prescriptions within 1 year prior to first plasma date

```{r}

## Function to generate variables on ICS/OCS under certain restriction

cs_vars_fnc <- function(restriction, varname_suffix) {
        cs_tmp <- eval(parse(text = str_c("med_cs[", restriction, ", ]")))
        cs_per_subj <- cs_tmp[, .(n_ics_rec = sum(ics, na.rm = T), 
                                  n_ocs_rec = sum(ocs, na.rm = T)), EMPI]
        cs_per_subj <- merge(cs_per_subj, cs_tmp[ics == 1, .(ics_closest_days_to1st = max(med_timegap_to1st_d)), EMPI], 
                             by = "EMPI", all.x = T)
        cs_per_subj <- merge(cs_per_subj, cs_tmp[ocs == 1, .(ocs_closest_days_to1st = max(med_timegap_to1st_d)), EMPI], 
                             by = "EMPI", all.x = T)
        setnames(cs_per_subj, c("EMPI", str_c(colnames(cs_per_subj)[-1], varname_suffix)))
        return(cs_per_subj)
}


## Generate # of ICS/OCS prescriptions within 5/2/1 year prior to last plasma date

dat <- pool[, .(EMPI, Biobank_Subject_ID, asthma_existence, asthma_ppv, first_plasma_date, last_plasma_date)]
ocs_ics_dat <- merge(dat, 
                     cs_vars_fnc(restriction = "med_timegap_to1st_d <= 0 & med_timegap_to1st_d >= -5", 
                     varname_suffix = "_5ybf_1stplasma"), 
                     by = "EMPI", all.x = T)
ocs_ics_dat <- merge(ocs_ics_dat, 
                     cs_vars_fnc(restriction = "med_timegap_to1st_d <= 0 & med_timegap_to1st_d >= -2", 
                     varname_suffix = "_2ybf_1stplasma"), 
                     by = "EMPI", all.x = T)
ocs_ics_dat <- merge(ocs_ics_dat, 
                     cs_vars_fnc(restriction = "med_timegap_to1st_d <= 0 & med_timegap_to1st_d >= -1", 
                     varname_suffix = "_1ybf_1stplasma"), 
                     by = "EMPI", all.x = T)
fwrite(ocs_ics_dat, file = here(res_dir, "ocs_ics_data_before_first_plasma_date_all_subjects.csv"))

```

Variable names:  

 - n_ics_rec: number of ICS prescription records (if NA means subject didn't have any ics/ocs data)
 - ics_closest_days_to1st: number of days between closest ics/ocs record and first plasma date within specified time frame
 - n_ocs_rec: number of OCS prescription records (if NA means subject didn't have any ics/ocs data) 
 - ocs_closest_days_to1st: number of days between closest ics/ocs record and first plasma date within specified time frame

Suffix (to specify time frame):

 - _5ybf_1stplasma: five years before first plasma collection date
 - _2ybf_1stplasma: two years before first plasma collection date
 - _1ybf_1stplasma: one year before first plasma collection date 

# Session info

```{r}

sessionInfo()

```
