---
title: "Significant features in PEGASUS global metabolomics"
author: "MHuang"
date: "12/4/2020"
output: 
  html_document: 
    toc: yes
    # toc_float: 
    #   collapsed: false
    #   smooth_scroll: false
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

## Packages

pkg <- c("tidyverse", "data.table", "here", "readxl", "gtools", "DT", "corrplot", "pheatmap", "RColorBrewer", 
         "viridis", "rstatix", "ggvenn")
for (p in pkg) {
        if (require(p, character.only = T)) {
                print(paste0(p, " loaded successfully"))
        } else {
                install.packages(p)
                require(p, character.only = T)
                print(paste0(p, " downloaded and loaded successfully"))
        }
}


## Paths

dat_dir <- "data/"
fig_dir <- "figures/"
res_dir <- "results/"
res_dir_hilic_pos <- "results/hilic_pos/"
res_dir_lipid_pos <- "results/lipid_pos/"
res_dir_lipid_neg <- "results/lipid_neg/"
res_dir_sph <- "results/sphingolipids/"
annot_dir <- "data/annotations/"


## File names

hilic_pos_dat_fname <- here(dat_dir, "hilic_pos_data_updated.RData")
lipid_pos_dat_fname <- here(dat_dir, "lipid_pos_data_updated.RData")
lipid_neg_dat_fname <- here(dat_dir, "lipid_neg_data_updated.RData")
res_name_list <- c("_overall_", "_female_", "_male_", "_ics_1y_y_", "_ics_1y_n_", "_eos_high_", "_eos_low_", 
                   "_bmi_high_", "_bmi_low_", "_allergy_y_", "_allergy_n_", "_neut_high_", "_neut_low_")

res_asth_hilic_pos_fname <- here(res_dir_hilic_pos, "res_compare_dif_asthmatics_hilic_pos.rds")
res_asth_lipid_pos_fname <- here(res_dir_lipid_pos, "res_compare_dif_asthmatics_lipid_pos.rds")
res_asth_lipid_neg_fname <- here(res_dir_lipid_neg, "res_compare_dif_asthmatics_lipid_neg.rds")
res_asth_sph_fname <- here(res_dir_sph, "res_compare_dif_asthmatics_sph.rds")

res_qt_hilic_pos_fname <- here(res_dir_hilic_pos, "res_qt_asthmatics_hilic_pos.rds")
res_qt_lipid_pos_fname <- here(res_dir_lipid_pos, "res_qt_asthmatics_lipid_pos.rds")
res_qt_lipid_neg_fname <- here(res_dir_lipid_neg, "res_qt_asthmatics_lipid_neg.rds")
res_qt_sph_fname <- here(res_dir_sph, "res_qt_asthmatics_sph.rds")

annot_fname <- here(annot_dir, "PEGASuS_Annotations_PeiZhang.xlsx")


## Hard-coded numbers

missing_thld <- 0.30
sig_digits <- 3
alpha_thld <- 0.05
z_95 <- qnorm(p = 0.975)
keep_cols <- c("subgroup", "feature", "tentative_annotation", "med_feat", "wo_ms2", "platform", "totaln", "casen", 
               "beta", "pval", "or", "lower95", "upper95", "pct_na_met", "cv", "expo_linreg", "beta_linreg", "pval_linreg")
num_cols <- c("beta", "pval", "fdr_bh", "or", "lower95", "upper95", "beta_linreg", "pval_linreg", "neg_log10pval")


## Function

save_pheatmap_png <- function(x, filename, width = 18, height = 18, units = "in", res = 300) {
        png(filename, width = width, height = height, units = units, res = res)
        grid::grid.newpage()
        grid::grid.draw(x$gtable)
        dev.off()
}

```

# 1. Asthma vs controls results 
## 1.1. HILIC positive results

```{r}

res_hilic_pos <- NULL
for (n in res_name_list) {
        eval(parse(text = str_c("res", n, "hilic_pos <- readRDS(file = here(res_dir_hilic_pos, 'res", 
                                n, "hilic_pos.rds'))")))
        eval(parse(text = str_c("res_hilic_pos <- rbind(res_hilic_pos, res", n, "hilic_pos", 
                                "[, ..keep_cols])")))
        eval(parse(text = str_c("rm(res", n, "hilic_pos)")))
}

nrow(res_hilic_pos[subgroup == "overall" & tentative_annotation != "Unknown" & pval < alpha_thld, ])
nrow(res_hilic_pos[subgroup == "overall" & tentative_annotation == "Unknown" & pval < alpha_thld, ])

annot_hilic_pos <- read_xlsx(annot_fname, sheet = "HILIC_POS") %>% as.data.table()
setnames(annot_hilic_pos, gsub(" ", "_", tolower(colnames(annot_hilic_pos))))
res_hilic_pos <- merge(res_hilic_pos, annot_hilic_pos[, .(feature, annotation, confidence_level, comment)], 
                       by = "feature", all.x = T)

res_asth_hilic_pos <- readRDS(res_asth_hilic_pos_fname)
res_asth_hilic_pos <- merge(res_asth_hilic_pos, annot_hilic_pos[, .(feature, annotation, confidence_level, comment)], 
                            by = "feature", all.x = T)

res_qt_hilic_pos <- readRDS(res_qt_hilic_pos_fname)
res_qt_hilic_pos <- merge(res_qt_hilic_pos, annot_hilic_pos[, .(feature, annotation, confidence_level, comment)], 
                          by = "feature", all.x = T)

```

## 1.2. Lipidomics positive results

```{r}

res_lipid_pos <- NULL
for (n in res_name_list) {
        eval(parse(text = str_c("res", n, "lipid_pos <- readRDS(file = here(res_dir_lipid_pos, 'res", 
                                n, "lipid_pos.rds'))")))
        eval(parse(text = str_c("res_lipid_pos <- rbind(res_lipid_pos, res", n, "lipid_pos", 
                                "[, ..keep_cols])")))
        eval(parse(text = str_c("rm(res", n, "lipid_pos)")))
}

nrow(res_lipid_pos[subgroup == "overall" & tentative_annotation != "Unknown" & pval < alpha_thld, ])
nrow(res_lipid_pos[subgroup == "overall" & tentative_annotation == "Unknown" & pval < alpha_thld, ])

annot_lipid_pos <- read_xlsx(annot_fname, sheet = "LIPIDOMICS_POS") %>% as.data.table()
setnames(annot_lipid_pos, gsub(" ", "_", tolower(colnames(annot_lipid_pos))))
res_lipid_pos <- merge(res_lipid_pos, annot_lipid_pos[, .(feature, annotation, confidence_level, comment)], 
                       by = "feature", all.x = T)

res_asth_lipid_pos <- readRDS(res_asth_lipid_pos_fname)
res_asth_lipid_pos <- merge(res_asth_lipid_pos, annot_lipid_pos[, .(feature, annotation, confidence_level, comment)], 
                            by = "feature", all.x = T)

res_qt_lipid_pos <- readRDS(res_qt_lipid_pos_fname)
res_qt_lipid_pos <- merge(res_qt_lipid_pos, annot_lipid_pos[, .(feature, annotation, confidence_level, comment)], 
                          by = "feature", all.x = T)

```

## 1.3. Lipidomics negative results

```{r}

res_lipid_neg <- NULL
for (n in res_name_list) {
        eval(parse(text = str_c("res", n, "lipid_neg <- readRDS(file = here(res_dir_lipid_neg, 'res", 
                                n, "lipid_neg.rds'))")))
        eval(parse(text = str_c("res_lipid_neg <- rbind(res_lipid_neg, res", n, "lipid_neg", 
                                "[, ..keep_cols])")))
        eval(parse(text = str_c("rm(res", n, "lipid_neg)")))
}

nrow(res_lipid_neg[subgroup == "overall" & tentative_annotation != "Unknown" & pval < alpha_thld, ])
nrow(res_lipid_neg[subgroup == "overall" & tentative_annotation == "Unknown" & pval < alpha_thld, ])

annot_lipid_neg <- read_xlsx(annot_fname, sheet = "LIPIDOMICS_NEG") %>% as.data.table()
setnames(annot_lipid_neg, "note", "confidence_level")
res_lipid_neg <- merge(res_lipid_neg, annot_lipid_neg[, .(feature, annotation, confidence_level)], 
                       by = "feature", all.x = T)
res_lipid_neg[, comment := NA]

res_asth_lipid_neg <- readRDS(res_asth_lipid_neg_fname)
res_asth_lipid_neg <- merge(res_asth_lipid_neg, annot_lipid_neg[, .(feature, annotation, confidence_level)], 
                            by = "feature", all.x = T)
res_asth_lipid_neg[, comment := NA]

res_qt_lipid_neg <- readRDS(res_qt_lipid_neg_fname)
res_qt_lipid_neg <- merge(res_qt_lipid_neg, annot_lipid_neg[, .(feature, annotation, confidence_level)], 
                          by = "feature", all.x = T)
res_qt_lipid_neg[, comment := NA]

```

## 1.4. Targeted sphingolipids results

```{r}

res_sph <- NULL
for (n in res_name_list) {
        eval(parse(text = str_c("res", n, "sph <- readRDS(file = here(res_dir_sph, 'res", 
                                n, "sph.rds'))")))
        eval(parse(text = str_c("res_sph <- rbind(res_sph, res", n, "sph", 
                                "[, ..keep_cols])")))
        eval(parse(text = str_c("rm(res", n, "sph)")))
}

nrow(res_sph[subgroup == "overall" & tentative_annotation != "Unknown" & pval < alpha_thld, ])
nrow(res_sph[subgroup == "overall" & tentative_annotation == "Unknown" & pval < alpha_thld, ])

res_sph[, ':='(annotation = NA, confidence_level = NA, comment = NA)]

res_asth_sph <- readRDS(res_asth_sph_fname)
res_asth_sph[, ':='(annotation = NA, confidence_level = NA, comment = NA)]

res_qt_sph <- readRDS(res_qt_sph_fname)
res_qt_sph[, ':='(annotation = NA, confidence_level = NA, comment = NA)]

```

## 1.5. Concatenate all results

```{r}

res_all <- rbind(res_hilic_pos, res_lipid_pos, res_lipid_neg, res_sph)
res_all[grepl("Montelukast|Mequitazine|Acetaminophen|Omeprazole|Tolmetin", annotation, ignore.case = T), 
        med_feat := 1][is.na(med_feat), med_feat := 0]
res_all[, fdr_bh := p.adjust(pval, method = "BH"), by = subgroup]
res_all[, ':='(feat_id = ifelse(is.na(annotation), str_c(feature, "_", tentative_annotation), 
                            str_c(feature, "_", annotation)), 
               med_feat = as.factor(med_feat), 
               subgroup = factor(subgroup, levels = c("overall", "female", "male", "ics_1y_prior_yes", 
                                                      "ics_1y_prior_no", "eos_high_300", "eos_low_300", 
                                                      "bmi_high_30", "bmi_low_30", "ige_high_env_allergy", 
                                                      "ige_norm_no_env_allergy", "neut_high_median", "neut_low_median")), 
               platform = as.factor(platform), 
               neg_log10pval = -log10(pval))]
setcolorder(res_all, c("platform", "feature", "annotation", "confidence_level", "comment", "subgroup", 
                       "totaln", "beta", "pval", "fdr_bh"))
saveRDS(res_all, file = here(res_dir, "res_all.rds"))

unique(res_all[platform == "HILIC-pos", .(subgroup, totaln, casen)])


## Features with p<0.05 in overall analyses

res_all[subgroup == "overall" & pval < alpha_thld, .N, .(platform)]

# res_all[subgroup == "overall" & pval < alpha_thld & !is.na(annotation), ][order(pval)] %>%
#         datatable(filter = "top") %>% formatSignif(columns = num_cols, digits = 3)
feat_overall_005 <- res_all[subgroup == "overall" & pval < alpha_thld & !is.na(annotation), feat_id]
length(unique(feat_overall_005)) == length(feat_overall_005); length(feat_overall_005)

feat_005 <- res_all[pval < alpha_thld & !is.na(annotation), unique(feat_id)]
length(unique(feat_005)) == length(feat_005); length(feat_005)
table(feat_005 %in% feat_overall_005) # same as those from overall analysis


## Features with p<0.01 in overall analyses

feat_overall_001 <- res_all[subgroup == "overall" & pval < alpha_thld/5 & !is.na(annotation), feat_id]
length(unique(feat_overall_001)) == length(feat_overall_001); length(feat_overall_001)

# fwrite(res_all[feat_id %in% feat_overall_005 & subgroup == "overall", ], 
#        file = here(res_dir, "features_significant_in_overall_nominal.csv"))

```

## 1.6. Features with p<0.05 in overall analysis 

> include features fdr<0.05 in subgroup analyses  

Subgroup analyses:  
 - by sex F/M  
 - by number of ICS prescriptions within 1 year prior to serum collection > 0 or not  
 - by eosinophil count closest to serum collection >= 300/uL or not  
 - by number of asthma exacerbations within 1 year prior to serum collection > 0 or not  
 - by BMI >= 30 or not  
 - by total/total IgE high or existence of environmental allergy 
 - by neutrophil count higher than median or not  

### 1.6.1. Results tables

```{r}

datatable(res_all[feat_id %in% feat_overall_005][order(subgroup, pval)], filter = "top") %>% 
        formatSignif(columns = num_cols, digits = 3)
# fwrite(res_all[feat_id %in% feat_overall_005][order(-subgroup, pval)], 
#        file = here(res_dir, "res_sig_sort_by_subgrp_pval.csv"))

# datatable(res_all[feat_id %in% feat_overall_005][order(tentative_annotation, subgroup)], 
#           filter = "top") %>% formatSignif(columns = num_cols, digits = 3)
# fwrite(res_all[feat_id %in% feat_overall_005][order(tentative_annotation, subgroup)], 
#        file = here(res_dir, "res_sig_sort_by_annot_subgrp.csv"))

# fwrite(res_all[feat_id %in% feat_overall_005 & pval < alpha_thld][order(tentative_annotation, pval)],
#        file = here(res_dir, "res_sig_sort_by_annot_pval_005.csv"))

```

### 1.6.2. Heatmaps (non-medication features)

Features with p<0.01 in overall analysis & features fdr<0.05 in subgroup analyses

```{r}

## -log10(pval)

pval_mat_001 <- dcast(res_all[feat_id %in% feat_overall_001 & med_feat == 0, ], feat_id ~ subgroup, 
                      value.var = "neg_log10pval") %>% as.data.frame()
rownames(pval_mat_001) <- pval_mat_001$feat_id
pval_mat_001$feat_id <- NULL
pheatmap(pval_mat_001, cluster_rows = T, cluster_cols = F, angle_col = 315, fontsize = 16, 
         main = "-log10(p-value)", color = inferno(10, direction = -1)) %>% 
        save_pheatmap_png(., here(fig_dir, "heatmap_neg_log10pval_001.png"), width = 21, height = 21)
pheatmap(subset(pval_mat_001, select = c("overall", "eos_high_300", "bmi_high_30", "ige_high_env_allergy")), 
         cluster_rows = T, cluster_cols = F, angle_col = 315, fontsize = 16, 
         main = "-log10(p-value)", color = inferno(10, direction = -1)) %>% 
        save_pheatmap_png(., here(fig_dir, "heatmap_neg_log10pval_001_subtype.png"), width = 12, height = 21)


pval_mat_005 <- dcast(res_all[feat_id %in% feat_overall_005 & med_feat == 0, ], feat_id ~ subgroup, 
                      value.var = "neg_log10pval") %>% as.data.frame()
rownames(pval_mat_005) <- pval_mat_005$feat_id
pval_mat_005$feat_id <- NULL
pheatmap(pval_mat_005, cluster_rows = T, cluster_cols = F, angle_col = 315, fontsize = 16, 
         main = "-log10(p-value)", color = inferno(10, direction = -1)) %>% 
        save_pheatmap_png(., here(fig_dir, "heatmap_neg_log10pval_005.png"), width = 21, height = 42)
pheatmap(subset(pval_mat_005, select = c("overall", "eos_high_300", "bmi_high_30", "ige_high_env_allergy")), 
         cluster_rows = T, cluster_cols = F, angle_col = 315, fontsize = 16, 
         main = "-log10(p-value)", color = inferno(10, direction = -1)) %>% 
        save_pheatmap_png(., here(fig_dir, "heatmap_neg_log10pval_005_subtype.png"), width = 12, height = 42)


# ## beta
# 
# beta_mat <- dcast(res_all[feat_id %in% c(feat_overall_001, feat_subgrp_fdr_005) & med_feat == 0, ], feat_id ~ subgroup, 
#                   value.var = "beta") 
# rownames(beta_mat) <- beta_mat$feat_id
# beta_mat$feat_id <- NULL
# pheatmap(beta_mat, cluster_rows = T, cluster_cols = F, 
#          main = "beta estimate") %>% 
#         save_pheatmap_png(., here(fig_dir, "heatmap_beta.png"))


## Venn Diagram

list(`overall` = res_all[subgroup == "overall" & pval < alpha_thld, feat_id], 
     `eos_high_300` = res_all[subgroup == "eos_high_300" & pval < alpha_thld, feat_id], 
     `bmi_high_30` = res_all[subgroup == "bmi_high_30" & pval < alpha_thld, feat_id], 
     `ige_high_env_allergy` = res_all[subgroup == "ige_high_env_allergy" & pval < alpha_thld, feat_id]) %>%
        ggvenn(stroke_size = 0.5, set_name_size = 5, text_size = 5, show_percentage = F) +  
               labs(title = "Number of features associated with asthma status (p<0.05)") + 
               theme(plot.title = element_text(hjust = 0.5, size = 15))
ggsave(file = here(fig_dir, "venn_overall_subtype_pval005.png"), width = 8, height = 8)


list(`overall` = res_all[subgroup == "overall" & fdr_bh < alpha_thld, feat_id], 
     `eos_high_300` = res_all[subgroup == "eos_high_300" & fdr_bh < alpha_thld, feat_id], 
     `bmi_high_30` = res_all[subgroup == "bmi_high_30" & fdr_bh < alpha_thld, feat_id], 
     `ige_high_env_allergy` = res_all[subgroup == "ige_high_env_allergy" & fdr_bh < alpha_thld, feat_id]) %>%
        ggvenn(stroke_size = 0.5, set_name_size = 5, text_size = 5, show_percentage = F) +  
               labs(title = "Number of features associated with asthma status (fdr<0.05)") + 
               theme(plot.title = element_text(hjust = 0.5, size = 15))
# ggsave(file = here(fig_dir, "venn_overall_subtype_fdr005.png"), width = 8, height = 8)

```

> Lipidomics nomenclature

 | Abbreviation | Lipid subclass | Main class | 
 | ------------ | -------------- | ---------- | 
 | Cer | Ceramide | Ceramides | 
 | CerP | Ceramide 1-phosphates | Ceramides | 
 | FAHFA | Fatty acid ester of hydroxyl fatty acid | Fatty esters | 
 | HexCer | Hexosylceramide | Neutral glycosphingolipids | 
 | Hex2Cer | Dihexosylceramide | Neutral glycosphingolipids | 
 | Hex3Cer | Trihexosylceramide | Neutral glycosphingolipids | 
 | LPC | Lysophophatidylcholine | Glycerophosphocholines | 
 | LPE | Lysophosphatidylethanolamine | Glycerophosphoethanolamines | 
 | LPI | Lysophosphatidylinositol | Glycerophosphoinositols | 
 | NAE | N-acyl ethanolamines | Fatty amides | 
 | TG | Triacylglycerol | Triradylglycerols | 
 | PA | Phosphatidic acid | Glycerophosphates | 
 | PC | Phosphatidylcholine | Glycerophosphocholines | 
 | PE | Phosphatidylethanolamine | Glycerophosphoethanolamines | 
 | PMeOH | Phosphatidylmethanol | Other Glycerophospholipids | 

### 1.6.3. Concatenate gender specific results for Priya
#### 1.6.3.1. Female

```{r, eval = F}

res_all_female_sig <- res_all[subgroup == "female" & pval < alpha_thld, ..keep_cols]
fwrite(res_all_female_sig[order(pval)], file = here(res_dir, "res_all_female_sig.csv"))

```

#### 1.6.3.2. Male

```{r, eval = F}

res_all_male_sig <- res_all[subgroup == "male" & pval < alpha_thld, ..keep_cols]
fwrite(res_all_male_sig[order(pval)], file = here(res_dir, "res_all_male_sig.csv"))

```

# 2. Within asthmatics comparison results
## 2.1. Binary comparisons

```{r}

res_asth <- rbind(res_asth_hilic_pos, res_asth_lipid_pos, res_asth_lipid_neg, res_asth_sph)
res_asth[grepl("Montelukast|Mequitazine|Acetaminophen|Omeprazole|Tolmetin", annotation, ignore.case = T), 
         med_feat := 1][is.na(med_feat), med_feat := 0]
res_asth[, fdr_bh := p.adjust(pval, method = "BH"), by = comparison]
res_asth[, ':='(feat_id = ifelse(is.na(annotation), str_c(feature, "_", tentative_annotation), 
                            str_c(feature, "_", annotation)), 
                med_feat = as.factor(med_feat), 
                comparison = factor(comparison, 
                                    levels = c("high_vs_low_eos_asthmatics", "obese_vs_not_asthmatics", 
                                               "allergy_vs_not_asthmatics", "female_vs_male_asthmatics", 
                                               "female_vs_male_controls")), 
                platform = as.factor(platform), 
                neg_log10pval = -log10(pval))]
setcolorder(res_asth, c("platform", "feature", "annotation", "confidence_level", "comment", "comparison", 
                        "beta", "pval", "fdr_bh"))
saveRDS(res_asth, file = here(res_dir, "res_asth.rds"))

datatable(res_asth[pval < alpha_thld & !(is.na(annotation) & tentative_annotation == "Unknown"), c(1:16)
                   ][order(comparison, pval)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "cv", "pct_na_met"), digits = 3)


## -log10(pval)

pval_mat_asth_001 <- dcast(res_asth[feat_id %in% feat_overall_001 & med_feat == 0, ], feat_id ~ comparison, 
                           value.var = "neg_log10pval") %>% as.data.frame()
rownames(pval_mat_asth_001) <- pval_mat_asth_001$feat_id
pval_mat_asth_001$feat_id <- NULL
pheatmap(pval_mat_asth_001, cluster_rows = T, cluster_cols = F, angle_col = 315, fontsize = 16, 
         main = "-log10(p-value)", color = inferno(10, direction = -1)) %>% 
        save_pheatmap_png(., here(fig_dir, "heatmap_neg_log10pval_001_compare_asthma_subtypes.png"), width = 12, height = 21)


pval_mat_asth_005 <- dcast(res_asth[feat_id %in% feat_overall_005 & med_feat == 0, ], feat_id ~ comparison, 
                           value.var = "neg_log10pval") %>% as.data.frame()
rownames(pval_mat_asth_005) <- pval_mat_asth_005$feat_id
pval_mat_asth_005$feat_id <- NULL
pheatmap(pval_mat_asth_005, cluster_rows = T, cluster_cols = F, angle_col = 315, fontsize = 16, 
         main = "-log10(p-value)", color = inferno(20, direction = -1)) %>% 
        save_pheatmap_png(., here(fig_dir, "heatmap_neg_log10pval_005_compare_asthma_subtypes.png"), width = 12, height = 42)


## Venn Diagram

list(`high_vs_low_eos` = res_asth[comparison == "high_vs_low_eos_asthmatics" & pval < alpha_thld, feat_id], 
     `obese_vs_not` = res_asth[comparison == "obese_vs_not_asthmatics" & pval < alpha_thld, feat_id], 
     `allergy_vs_not` = res_asth[comparison == "allergy_vs_not_asthmatics" & pval < alpha_thld, feat_id], 
     `female_vs_male` = res_asth[comparison == "female_vs_male_asthmatics" & pval < alpha_thld, feat_id]) %>%
        ggvenn(stroke_size = 0.5, set_name_size = 5, text_size = 5, show_percentage = F) +  
               labs(title = "Number of features associated with binary outcomes in asthmatics (p<0.05)") + 
               theme(plot.title = element_text(hjust = 0.5, size = 15))
ggsave(file = here(fig_dir, "venn_bin_outc_asthmatics_pval005.png"), width = 8, height = 8)

```

## 2.2. Quantitative traits

```{r}

res_qt <- rbind(res_qt_hilic_pos, res_qt_lipid_pos, res_qt_lipid_neg, res_qt_sph)
res_qt[grepl("Montelukast|Mequitazine|Acetaminophen|Omeprazole|Tolmetin", annotation, ignore.case = T), 
       med_feat := 1][is.na(med_feat), med_feat := 0]
res_qt[, fdr_bh := p.adjust(pval, method = "BH"), by = comparison]
res_qt[, ':='(feat_id = ifelse(is.na(annotation), str_c(feature, "_", tentative_annotation), 
                               str_c(feature, "_", annotation)), 
              med_feat = as.factor(med_feat), 
              comparison = factor(comparison, 
                                  levels = c("eos_continous", "neut_continous", "fev1_continous", "fvc_continous")), 
              platform = as.factor(platform), 
              neg_log10pval = -log10(pval))]
setcolorder(res_qt, c("platform", "feature", "annotation", "confidence_level", "comment", "comparison", 
                      "beta", "pval", "fdr_bh"))
saveRDS(res_qt, file = here(res_dir, "res_qt.rds"))

datatable(res_qt[pval < alpha_thld & !(is.na(annotation) & tentative_annotation == "Unknown"), c(1:17)
                 ][order(comparison, pval)], filter = "top") %>% 
        formatSignif(columns = c("beta", "pval", "fdr_bh", "cv", "pct_na_met"), digits = 3)


## -log10(pval)

pval_mat_qt_001 <- dcast(res_qt[feat_id %in% feat_overall_001 & med_feat == 0, ], feat_id ~ comparison, 
                                value.var = "neg_log10pval") %>% as.data.frame()
rownames(pval_mat_qt_001) <- pval_mat_qt_001$feat_id
pval_mat_qt_001$feat_id <- NULL
pheatmap(pval_mat_qt_001, cluster_rows = T, cluster_cols = F, angle_col = 315, fontsize = 16, 
         main = "-log10(p-value)", color = inferno(6, begin = 1, end = 0.5, direction = 1)) %>% 
  save_pheatmap_png(., here(fig_dir, "heatmap_neg_log10pval_001_qt.png"), width = 9, height = 21)


pval_mat_qt_005 <- dcast(res_qt[feat_id %in% feat_overall_005 & med_feat == 0, ], feat_id ~ comparison, 
                                value.var = "neg_log10pval") %>% as.data.frame()
rownames(pval_mat_qt_005) <- pval_mat_qt_005$feat_id
pval_mat_qt_005$feat_id <- NULL
pheatmap(pval_mat_qt_005, cluster_rows = T, cluster_cols = F, angle_col = 315, fontsize = 16, 
         main = "-log10(p-value)", color = inferno(12, begin = 1, end = 0.3, direction = 1)) %>% 
  save_pheatmap_png(., here(fig_dir, "heatmap_neg_log10pval_005_qt.png"), width = 9, height = 42)


## Venn Diagram

list(`eos_continous` = res_qt[comparison == "eos_continous" & pval < alpha_thld, feat_id], 
     `neut_continous` = res_qt[comparison == "neut_continous" & pval < alpha_thld, feat_id], 
     `fev1_continous` = res_qt[comparison == "fev1_continous" & pval < alpha_thld, feat_id], 
     `fvc_continous` = res_qt[comparison == "fvc_continous" & pval < alpha_thld, feat_id]) %>%
        ggvenn(stroke_size = 0.5, set_name_size = 5, text_size = 5, show_percentage = F) +  
               labs(title = "Number of features associated with quantitative outcomes in asthmatics (p<0.05)") + 
               theme(plot.title = element_text(hjust = 0.5, size = 15))
ggsave(file = here(fig_dir, "venn_qt_outc_asthmatics_pval005.png"), width = 8, height = 8)

list(`eos_continous` = res_qt[comparison == "eos_continous" & fdr_bh < alpha_thld*4, feat_id], 
     `neut_continous` = res_qt[comparison == "neut_continous" & fdr_bh < alpha_thld*4, feat_id], 
     `fev1_continous` = res_qt[comparison == "fev1_continous" & fdr_bh < alpha_thld*4, feat_id], 
     `fvc_continous` = res_qt[comparison == "fvc_continous" & fdr_bh < alpha_thld*4, feat_id]) %>%
        ggvenn(stroke_size = 0.5, set_name_size = 5, text_size = 5, show_percentage = F) +  
               labs(title = "Number of features associated with quantitative outcomes in asthmatics (fdr<0.2)") + 
               theme(plot.title = element_text(hjust = 0.5, size = 15))
# ggsave(file = here(fig_dir, "venn_qt_outc_asthmatics_fdr02.png"), width = 8, height = 8)

```

# Session info

```{r}

sessionInfo()

```
